### 2020-07-17 report/report.tex
beginning of the report

\documentclass[a4paper, 10pt]{article}

\author{Guillaume Cluzel}
\title{Toward a secure TCP/IP stack --- Technical report}

\usepackage{amsmath,amsfonts}

\usepackage{listings}

\lstset{
    basicstyle=\ttfamily
}

\usepackage{tikz}
\usetikzlibrary{calc,backgrounds,shapes}
\usetikzlibrary{arrows.meta, shapes.callouts}

\tikzset{global scale/.style={
    scale=#1,
    every node/.style={scale=#1}
  }
}

\usepackage{hyperref}

\begin{document}

    \maketitle

    \section{Introduction}

    TCP is a widely used network protocol to communicate in the Internet as
    it is used by the HTTP protocol, the FTP protocol, and so many others.
    Ensuring the safety of the TCP/IP stack is essential for the safety of
    a lot of machines. While a lot of work has been done on higher level
    protocols, such as SSL/TLS which is built on top of TCP and designed
    to provide a security layer to TCP that have been the subject of lots of
    works like miTLS \cite{}, nothing has been done to make a secure TCP
    implementation. Nonetheless, TLS security can only be ensured if the underlying
    TCP implementation is free of bugs.

    CycloneTCP is a library developped by Oryx Embedded for embedded platforms.
    A large number of platforms are supported, and the library can be used
    with a dozen of OS. The code of the library is written in C and contains
    plenty protocols from low-level network layers like IPv4/IPv6, to transport layer
    and application layer like DHCP or HTTP. In particular an implementation of
    the TCP protocol is provided.

    SPARK 2014 is a programming language designed as a subset of Ada that helps in
    making high reliability sofwares by providing powerful static analysis methods.
    SPARK is able to detect uninitialized variables with control flow analysis,
    and can ensure the absence of run-time errors,
    but also, based on SMT-solver, functionnal behavior can be specified for every
    functions and mathematically proved.

    The aim of this work was to translate some parts of the CycloneTCP stack
    in Ada/SPARK to improve the safety of the code. In particular we focused
    on the TCP protocol to ensure that the norm is respected in this work.

    This report is a synthesis of the challenge of verifying a such protocol
    and the solution I found to go toward a secure protocol.

    \section{The TCP protocol}

    \subsection{Why verifying a TCP stack?}

    As mentionned in the introduction, TCP protocol is the base of a lot of other protocols,
    in particular SSL/TLS.

    \subsection{Sockets}

    A socket is nothing more than a data structures that contains the connection
    information, like for example the local or the remote IP address as well as
    the state of the connection. In Ada, a socket will be represented by a pointer
    to a record type:

    \begin{lstlisting}[language=Ada]
type Socket_Struct is record
    S_Descriptor     : Sock_Descriptor;
    S_Type           : Socket_Type;
    S_Protocol       : Socket_Protocol;
    S_Net_Interface  : System.Address;
    S_localIpAddr    : IpAddr;
    S_Local_Port     : Port;
    S_Remote_Ip_Addr : IpAddr;
    S_Remote_Port    : Port;
    S_Timeout        : Systime;
    State            : Tcp_State;
    -- Other fields
end record;
type Socket is access Socket_Struct;
    \end{lstlisting}

    A socket is also the structure that is manipulated by the user as an opaque structure,
    through an interface to perform operations on it and perform operations on the
    global environment by sending messages in the network for example.

    \subsection{Presentation of the protocol}

    TCP is defined by the norm RFC 793 \cite{} in a high level language.
    TCP is a realiable, ordered and error-checked connection oriented protocol.
    It means that two computers have to open a connection first before sending data.
    The connection is closed once all the data has been transmitted.
    The flow of a connection (if no error happens) contains three main steps:
    \begin{enumerate}
        \item Opening the connection,
        \item Sending \& receiving the data,
        \item Closing the connection.
    \end{enumerate}

    This mechanism is decribed by a state machine.

    \begin{figure}[p]
        \begin{tikzpicture}[>=latex, global scale=.75]

        %
        % Styles for states, and state edges
        %
        \tikzstyle{state} = [draw, very thick, fill=white, rectangle, minimum height=3em, minimum width=7em, node distance=8em, font={\sffamily\bfseries}]
        \tikzstyle{stateEdgePortion} = [black,thick];
        \tikzstyle{stateEdge} = [stateEdgePortion,->];
        \tikzstyle{edgeLabel} = [pos=0.5, text centered, font={\sffamily\small}];

        %
        % Position States
        %
        \node[state, name=closedStart] {CLOSED};
        \node[state, name=listen, below of=closedStart] {LISTEN};
        \node[state, name=synSent, below of=listen, right of=listen, xshift=8em] {SYN\_SENT};
        \node[state, name=synRcvd, below of=listen, left of=listen, xshift=-8em] {SYN\_RCVD};
        \node[state, name=established, below of=listen, node distance=14em] {ESTABLISHED};
        \node[state, name=finWait1, below of=established, left of=established, node distance=7em, xshift=-9em] {FIN\_WAIT\_1};
        \node[state, name=finWait2, below of=finWait1] {FIN\_WAIT\_2};
        \node[state, name=closeWait, below of=established, right of=established, node distance=7em, xshift=9em] {CLOSE\_WAIT};
        \node[state, name=closing, below of=established, node distance=14em] {CLOSING};
        \node[state, name=lastAck, below of=closeWait] {LAST\_ACK};
        \node[state, name=timeWait, below of=closing] {TIME\_WAIT};

        \node at ($(lastAck) + (0,-3)$) {Source: \url{http://www.texample.net}};

        %
        % Connect States via edges
        %
        \draw ($(closedStart.south) + (-.5em,0)$) 
        edge[stateEdge] node[edgeLabel, xshift=-3em]{\emph{Passive open}} 
        ($(listen.north) + (-.5em,0)$); 
        \draw ($(listen.north) + (.5em,0)$) 
        edge[stateEdge] node[edgeLabel, xshift=2em]{\emph{Close}} 
        ($(closedStart.south) + (.5em,0)$);

        \draw ($(listen.south) + (-1em,0)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-2em, yshift=1em]{SYN/SYN + ACK} 
        ($(synRcvd.east) + (0,1em)$);
        \draw ($(listen.south) + (1em,0)$) 
        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=1em, yshift=1em]{\emph{Send}/SYN} 
        ($(synSent.west) + (0,1em)$);

        \draw ($(synRcvd.north) + (.5em,0)$) 
        edge[stateEdge, bend left=45] node[edgeLabel,xshift=-4em]{\emph{Timeout}/RST} 
        ($(closedStart.west) + (0,-.5em)$);

        \draw ($(synSent.north) + (-.5em,0)$) 
        edge[stateEdge, bend right=45] node[edgeLabel,xshift=-1em, yshift=-1em]{\emph{Close}} 
        ($(closedStart.east) + (0,-.5em)$);
        \draw ($(closedStart.east) + (0,.5em)$) 
        edge[stateEdge, bend left=45] node[edgeLabel,xshift=4em]{\emph{Active open}/SYN} 
        ($(synSent.north) + (.5em,0)$);

        \draw (synSent.west) 
        edge[stateEdge] node[edgeLabel, yshift=1em]{SYN/SYN + ACK} 
        (synRcvd.east);
        \draw (synRcvd) 
        edge[stateEdge] node[edgeLabel, xshift=-2.5em]{\emph{Close}/FIN} 
        (finWait1);

        \draw ($(synRcvd.east) + (0,-1em)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1em, yshift=-1em]{ACK} 
        ($(established.north) + (-1em,0)$);
        \draw ($(synSent.west) + (0,-1em)$) 
        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=3em, yshift=-1em]{SYN + ACK/ACK} 
        ($(established.north) + (1em,0)$);

        \draw ($(established.south) + (-1em,0)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1em, yshift=1em]{\emph{Close}/FIN} 
        ($(finWait1.east) + (0,.5em)$);
        \draw ($(established.south) + (1em,0)$) 
        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=1em, yshift=1em]{FIN/ACK} 
        ($(closeWait.west) + (0,1em)$);

        \draw (finWait1.south) 
        edge[stateEdge] node[edgeLabel, xshift=-2em]{ACK} 
        (finWait2.north);
        \draw ($(finWait1.east) + (0,-.5em)$) 
        edge[stateEdge, bend left=22.5] node[edgeLabel, yshift=1em]{ACK} 
        (closing.north);
        \draw (finWait1.south east) 
        edge[stateEdge] node[edgeLabel, xshift=1em, yshift=2em, text width=3em]{FIN + ACK/ACK} 
        (timeWait.north west);

        \draw (finWait2.south) 
        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=-2em, yshift=-1em]{FIN/ACK} 
        (timeWait.west);

        \draw (closing) 
        edge[stateEdge] node[edgeLabel, xshift=-2em]{ACK} 
        (timeWait);

        \draw (closeWait) 
        edge[stateEdge] node[edgeLabel,xshift=2.5em]{\emph{Close}/FIN} 
        (lastAck);

        %
        % Connect lastAck to closed is slightly more complicated
        % no direct line-of-sight, so we need to take the scenic route
        %
        \coordinate (lastAck2ClosedA) at ($(lastAck.east) + (2em,0)$);
        \coordinate (lastAck2ClosedB) at ($(closedStart.north -| lastAck.east) + (2em,1em)$);
        \coordinate (lastAck2ClosedC) at ($(closedStart.north) + (0.5em,1em)$);
        \draw (lastAck.east) edge[stateEdgePortion] (lastAck2ClosedA);
        \draw (lastAck2ClosedA) edge[stateEdgePortion] node[edgeLabel,xshift=-1.5em, yshift=-4em]{ACK} (lastAck2ClosedB);
        \draw (lastAck2ClosedB) edge[stateEdgePortion] (lastAck2ClosedC);
        \draw (lastAck2ClosedC) edge[stateEdge] ($(closedStart.north) + (0.5em,0)$);

        %
        % likewise for timeWait to closed
        %
        \coordinate (timeWait2ClosedA) at ($(timeWait.south) + (0,-1em)$);
        \coordinate (timeWait2ClosedB) at ($(timeWait.south -| finWait2.west) + (-2em,-1em)$);
        \coordinate (timeWait2ClosedC) at ($(closedStart.north -| finWait2.west) + (-2em,1em)$);
        \coordinate (timeWait2ClosedD) at ($(closedStart.north) + (-0.5em,1em)$);
        \draw (timeWait.south) edge[stateEdgePortion] (timeWait2ClosedA);
        \draw (timeWait2ClosedA) edge[stateEdgePortion] (timeWait2ClosedB);
        \draw (timeWait2ClosedB) edge[stateEdgePortion] (timeWait2ClosedC);
        \draw (timeWait2ClosedC) edge[stateEdgePortion] 
        node[edgeLabel, text width=12.25em, yshift=1.5em]{\emph{Timeout after two maximum segment lifetimes (2*MSL)}} 
        (timeWait2ClosedD);
        \draw (timeWait2ClosedD) edge[stateEdge] ($(closedStart.north) + (-0.5em,0)$);

        % draw dotted lines around passive and active closes
        \begin{pgfonlayer}{background}
            \draw [join=round,black,dotted] ($(closeWait.north west) + (-1em, -1em)$) rectangle ($(lastAck.south east) + (1em, 1em)$);
            \draw [join=round,black,dotted] ($(finWait1.north west) + (-1em, -1em)$) rectangle ($(timeWait.south east) + (1em, 1em)$);
        \end{pgfonlayer}

        \end{tikzpicture}
        \caption{TCP automaton.}
        \label{fig:TCPAutomaton}
    \end{figure}

    As it can be seen in the state machine, there exists different states that are
    subject to change when a message is segment is sent or received or when an action
    is done by the user. A label $A/B$ on an edge refers to a stimuli performed on the socket
    (the receiption of a message, a user action...) for $A$ and a flag sent for $B$.
    More precisely, for $A$, a text in italic shape refers to an action perform by
    the user (a call of a user function) or when it's a timeout, by a timer and a ACK,
    SYN or FIN refers to a flag contained in a received message. The flags in $B$ refer
    to the flags sent in response of the stimuli $A$.  Without giving
    unnecessary details on the format of a TCP header, we can describe the different
    flags that can be contained in a TCP header:
    \begin{description}
        \item[ACK] Acknowledgment field significant. The last message received by the sender
        is acknowledge. A field of the header contains the number of this segment.
        \item[SYN] Synchronize sequence number. This flag is sent in order to establish
        a connection.
        \item[FIN] No more data from sender. This flag is sent in order to close the
        connection.
    \end{description}

    Different states can be taken by a socket during its lifetime depending on the state
    of the connection, and also depending on the state of the remote TCP state. Each
    state has its own signification:
    \begin{description}
        \item[CLOSED] represents no connection state at all.
        \item[LISTEN] represents waiting for a connection request from any remote
        TCP and port.
        \item[SYN-SENT] represents waiting for a matching connection request
        after having sent a connection request.
        \item[SYN-RECEIVED] represents waiting for a confirming connection
        request acknowledgment after having both received and sent a
        connection request.
        \item[ESTABLISHED] represents an open connection, data received can be
        delivered to the user. The normal state for the data transfer phase
        of the connection.
        \item[FIN-WAIT-1] represents waiting for a connection termination request
        from the remote TCP, or an acknowledgment of the connection
        termination request previously sent.
        \item[FIN-WAIT-2] represents waiting for a connection termination request
        from the remote TCP.
        \item[CLOSE-WAIT] represents waiting for a connection termination request
        from the local user.
        \item[CLOSING] represents waiting for a connection termination request
        acknowledgment from the remote TCP.
        \item[LAST-ACK] represents waiting for an acknowledgment of the
        connection termination request previously sent to the remote TCP
        (which includes an acknowledgment of its connection termination
        request).
        \item[TIME-WAIT] represents waiting for enough time to pass to be sure
        the remote TCP received the acknowledgment of its connection
        termination request.
    \end{description}

    For more details on how two TCP can establish a connection, the reader can refer
    to the TCP norm where pleinty of scenario are described and explained, and more
    details are given on the state and segments that can be sent.

    \subsubsection{Multithreading in the implementation}

    Different threads can interact with a socket. The TCP norm gives an example with
    three threads: one for the \emph{user calls}, one for the \emph{arriving segments}
    and one for the \emph{timers}. This design has been adopted in CycloneTCP.

    \paragraph{User calls}
    User calls refers to functions that can be called by the user to control the connection,
    send, or received data: OPEN, CLOSE, ABORT, SEND, RECEIVED. Transition between state of
    the connection can happen during the call of these function since there are made to control
    the connection.

    \paragraph{Arriving segments}
    In this threads, the received segments are processed. The corresponding messages are sent back.
    Transition between state can happen here. For example, this is the case when a message containing
    a SYN is received when the socket is in the LISTEN state. It will automatically send a SYN + ACK in
    response, and change its state for SYN-RECEIVED.

    \paragraph{Timers}
    The timers control the timeouts like the retransmission timeout to retransmit a message, or
    the time-wait timeout to close the connection after an amont of time. Then, transition can also
    happen in this threads.

    To sum up, a single socket can be seen as a shared structure over multiples threads, that can all
    manipulate it and change its fields. The schema illustrates that.

    Only one thread can perform an operation on the socket at the same time. The access to the fields of
    the socket are protected by mutex.

    Two threads can communicate synchronously or asynchronously together. The synchronous communications are based on
    the interface provided by the OS, in particular with events.

    \begin{tikzpicture}
        \node [draw, ellipse] (S) at (2, 2) {Socket};
        \node [draw, text width=2.5cm, anchor=north] (T1) at (-2,0) {\textbf{TIMER TASK} \\
                                                Timeout when need};
        \node [draw, text width=2.5cm, anchor=north] (T2) at (2,0) {\textbf{USER TASK} \\
                                                Controling the connection on the user side};
        \node [draw, text width=2.5cm, anchor=north] (T3) at (6, 0) {\textbf{RECEIVED TASK} \\
                                                Processing received messages};
        \draw[->] (T1) -- (S);
        \draw[->] (T2) -- (S);
        \draw[->] (T3) -- (S);
        \draw[<->, dashed] (T1) -- (T2);
        \draw[<->, dashed] (T2) -- (T3);

        \draw [->] (4.3,2) -- (4.9,2);
        \node [anchor=west] at (5,2) {Operation};
        \draw[<->, dashed] (4.3,1.3) -- (4.9,1.3);
        \node [anchor=west, text width=2.7cm] at (5, 1.3) {Synchronous communications};
    \end{tikzpicture}

    \section{Challenges}

    \subsection{Main focus for the verification}

    The TCP norm defines a lot of behavior that could be verified by automatic proof.
    Here is a non exhaustive list of what could be done:
    \begin{itemize}
        \item Verification of the transitions between the states.
        \item Integrity of the messages sent, \textit{i.e.} we could check that the sent
        message contains the correct flag and the correct sequence and acknowledgement numbers.
        \item Integrity of the messages received, \textit{i.e.} we could check that the
        messages received are correctly processed in regard of the flags they contain.
        \item Functionnal correctness of the user functions in regard to their specifications.
    \end{itemize}

    Some choices have had to be done in regard the short time devolved to the internship, and
    not every aspect of the TCP norm has been prooved. However, some aspects have been selected
    for the challenge or the importance they represent. A list of these aspects is presented
    here and the choices will be motivated in the next subsections. First of all, we have wanted
    to improve the interface of the functions. The other point was to improved the safety of the
    protocol itself.

    \subsubsection{Improve functions interface}

    One of the main problems pointed by the main writter of the CycloneTCP library is that users
    do not correctly used the API supplied. As a result, some calls of
    functions can be incorrect, either because the arguments are incorrect, or because the
    return code of the previous functions has not been checked, which drive to an
    incorrect code.

    \subsubsection{Respect to the protocol}

    To ensure the safety of the library, it is necessary to ensure that the functions really
    do what they are supposed to do. It is a part of the job that has been done. More
    especially we have wanted to checked if the transition machine is respected by all the
    user functions. We have also tried to ensure the correctness of the function called
    regarding to the TCP state of the socket before the call to respect the specifications
    given in the norm.

    \subsection{Technical problems encountered}

    The multiples tasks and the changes that can happen at every moment in one
    task or another make the verification hard. Multiples interactions exist
    between the multiples threads: synchronous and asynchronous.
    All these interraction have to be considered if we want to write correct contracts
    in regard of what could be done in other threads.

    As SPARK does not have a native mode to deal with concurrency, we have to
    model these different interactions by hands by writing assertions that
    modify the state of the socket as another thread could do it. In particular
    the problem has been encountered when we tried to write contracts for the correctness
    of the user functions. More details and explainations will be given in the following
    section, where we will investigate the solution found to these technical problems.

    \section{Solutions found}

    \subsection{Order to call the functions}

    This part mainly focussed on the high level user functions, the socket functions.
    The socket functions are the one manipulated by the user to perform operations on
    the network. They are located in the files \texttt{socket\_interface.ad(b|s)}.

    It is clear that there exist an order on which the functions have to be called.
    This order can be determined by the norm, and must respect the order given by the
    graph in figure \ref{fig:TCPAutomaton}.
    Then, we want post- and pre-conditions to model a partial order on the calls of the function.
    If two functions $f_1$ and $f_2$ are ordered such that $f_1 \preceq f_2$ where $\preceq$ is
    a relation over the oder of which functions have to be called, then we want that the
    post-condition of $f_1$ implies the precondition of $f_2$.

    We will give an example over functions \lstinline[language=Ada]{Socket_Connect} and
    \lstinline[language=Ada]{Socket_Send}. The function \lstinline[language=Ada]{Socket_Connect}
    try to connect to a distant TCP. If the connection succeed, the remote IP address of the distant
    TCP is set in the field \lstinline[language=Ada]{S_Remote_Ip_Addr} of the \lstinline[language=Ada]{Sock}
    structure. When a user call the function \lstinline[language=Ada]{Socket_Send}, we want to ensure that
    the connection has already been established. This is why a precondition of this function is
    \lstinline[language=Ada]{Is_Initialized_Ip(Sock.S_Remote_Ip_Addr)}.

    \begin{figure}
        \begin{lstlisting}[language=Ada]
procedure Socket_Connect
    (Sock           : in out Not_Null_Socket;
     Remote_Ip_Addr : in     IpAddr;
     Remote_Port    : in     Port;
     Error          :    out Error_T)
  with
   Pre => Is_Initialized_Ip (Remote_Ip_Addr),
   Contract_Cases => (
     Sock.S_Type = SOCKET_TYPE_STREAM =>
       (if Error = NO_ERROR then
         Sock.S_Type = Sock.S_Type'Old and then
         Sock.S_Protocol = Sock.S_Protocol'Old and then
         Is_Initialized_Ip (Sock.S_localIpAddr) and then
         Sock.S_Local_Port = Sock.S_Local_Port'Old and then
         Sock.S_Remote_Ip_Addr = Remote_Ip_Addr and then
         Sock.S_Remote_Port = Remote_Port and then
         Sock.State = TCP_STATE_ESTABLISHED
       else
         Sock.S_Type = Sock.S_Type'Old and then
         Sock.S_Protocol = Sock.S_Protocol'Old)
     others => True)

procedure Socket_Send
    (Sock    : in out Not_Null_Socket;
     Data    : in     Send_Buffer;
     Written :    out Natural;
     Flags   :        Socket_Flags;
     Error   :    out Error_T)
  with
   Pre  =>
     Is_Initialized_Ip(Sock.S_Remote_Ip_Addr)
        \end{lstlisting}
        \caption{An example of how functions calls can be ordered by Pre- and Post-conditions.}
        \label{code:FunOrdered}
    \end{figure}

    \subsection{Check of return code}

    An observation made by Clément Zeller, the main programmer of the library is that customers
    do not always think to check the return code of the socket user functions, that can fail.
    If the return code is not checked, some assumption on the result cannot be done.

    The post-conditions as we have written them in SPARK, ensure that the return code is checked
    before continuating processing. Figure \ref{code:FunOrdered} shows this mechanism for the
    procedure \lstinline[language=Ada]{Socket_Connect}. The post-condition filters the cases
    where \lstinline{Error} is \lstinline{NO_ERROR} and if \lstinline{Error} takes another value.
    As a result, a user who would write an incorrect code such as
    \begin{lstlisting}[language=Ada]
    Socket_Connect (Sock, Remote_Ip_Addr, Port, Error);
    Socket_Send (Sock, Data, Written, Flags, Error);
    \end{lstlisting}
    would be warned by gnatprove with a message such as
    \begin{verbatim}
    medium: precondition might fail.
    \end{verbatim}
    whereas the following code is correct
    \begin{lstlisting}[language=Ada]
    Socket_Connect (Sock, Remote_Ip_Addr, Port, Error);
    if Error /= NO_ERROR then
        return;
    end if;
    Socket_Send (Sock, Data, Written, Flags, Error);
    \end{lstlisting}

    % \subsection{Liberties taken with the norm}

    % For some operations in the norm, if it is called when the socket in a particular
    % state, the function is assumed to return an error (see page 56 for an example of that).
    % I have made the choice to directly discrimine this choice in the pre condition
    % to allow user to check if its call is legal or not.

    \subsection{Verification of the state machine}

    The aim of this part is to explain how we have done to check that the transitions done
    in a function respect the TCP automaton. In particular, we were interested in verifying
    the high level user functions. These functions are not the one in which most of the
    transitions are done. Even, the verification of these functions is important in order
    to guarantee the safety of the whole library. The functions of interest are located in
    the file \texttt{tcp\_interface.ad(b|s)}.

    \subsubsection{Concurrency: asynchronous mechanism}

    To ensure the safety of our library, we need to consider all what can happen everywhere in
    the library. All our functions are protected by mutex, which means that only one operation
    can be performed at the same time on a socket. Interractions must be considered at two
    locations: between the function calls, when the mutex is released and during the function
    call, when the program waits for an event. We will explore in the two next sections how we
    have dealt with this two different concurrent mechanisms.

    Before everything else, let's see where the interference mainly comes from, and how we have
    done to ensure that all have been considered.

    \subsubsection{Process segments functions}

    Almost all the transitions are done in the file \texttt{tcp\_fsm.c}. This file has not been
    translated in SPARK by a lack of time. Then we do not have strong guaranties on its functions
    behavior. However, for formal verification we need to know what is done in the functions of
    this file, because a wrong contract can make all the verification wrong.

    The file \texttt{tcp\_fsm.c} is in charge of processing the incoming segments. The reader is
    strongly encouraged to have a look at this file. The main function of this file
    \texttt{tcpProcessSegment} look for the socket corresponding to the received segment, and then
    according to the TCP state of the socket, process the segment as expected by calling one of
    the function \texttt{tcpState<StateName>}. The functions familly \texttt{tcpState<StateName>}
    check the information contained in the segment and can perform change of state in the socket
    struct depending on the flags received.

    In these functions, change of state can be performed as mentionned, through the function
    \texttt{tcpChangeState}. To summarize what is done when a message is received, rather than
    reading the code and locate all the calls to functions \texttt{tcpChangeState}, we have used
    Klee\footnote{\url{https://klee.github.io}}. It helped to find behavior that had not been
    imagined at first. Anyway, we cannot state that all the work done with klee is complete, and
    we would need to rewrite this part in SPARK to have a code formally proved.

    All what is related to klee is present in the folder \texttt{klee/} and can be compiled and run
    thanks to the makefile to reproduce the results. Roughtly, what is done is creating a random
    incomming message, put the socket in the desired state and call the desired function to see what
    final states can be obtain.

    Finally, all the result found by klee has been reported in the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
    This function is essential for all the verification of concurrent parts.

    % The functions used to process the segments are written in C and we have not changed
    % them. We have just extracted some information, like the changes of state that can be
    % produced by the reception of a message.

    % To ensure the correctness of the extracted information, we have used Klee, a symbolic
    % virtual machine. It allows to consider all the paths taken in the function and check
    % if the contracts written in SPARK respects the C implementation.
    % \textbf{\itshape Show the example where I found an error.}

    % A goal to achieve would be to rewrite this part in SPARK to have more guaranties.
    % However, Klee is a tool that helps to find errors.

    \subsubsection{User functions}

    Code written in SPARK to check if the transitions are coherent inside the user functions.

    \subsubsection{Concurrency}

    Modeling the concurrency and the interleaving semantic has been the hardest work.
    The held solution is to have "ghost" code that model the different transition that
    might have happened when the mutex is released.

    The interleaving semantic is also considered for synchronous events, when the code
    wait for a particular event.

    If we consider the state transition given by the code that receive the message, we
    are driven to consider a transition relation, denotes as $\rightarrow$. This relation
    is reflexive because a received message can be drop if it doesn't contain useful
    information. Then, for all TCP state $S$, $S \rightarrow S$. As multiples messages
    can be received, the transitive closure of $\rightarrow$ is considered:
    $\rightarrow^* = \bigcup_{n\in\mathbb{N}^*} \rightarrow^n$ which can be simplified to:
    \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
    since every state is only reachable by a state at a distance at most of 3, which make
    the calculation of the transition closure easier by SPARK once we have $\rightarrow$ by
    unrolling loops.

\end{document}

### 2020-07-19 report/report.tex
Continue report

@@
 
+\usepackage[ruled]{algorithm2e}
+
 \begin{document}
@@
     \maketitle
+
+    \tableofcontents
 
@@
 
-    \subsubsection{Concurrency: asynchronous mechanism}
+    \subsubsection{Overview of the concurrency challenge}
 
@@
     Finally, all the result found by klee has been reported in the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
-    This function is essential for all the verification of concurrent parts.
+    This function is essential for all the verification of concurrent parts and is reused everywhere
+    the concurrency is involved to compute the results.
 
@@
 
-    \subsubsection{User functions}
+    \subsubsection{Concurrency: asynchronous changes of state}
 
-    Code written in SPARK to check if the transitions are coherent inside the user functions.
+    Between the user functions calls, segment can be received and these receptions can drive to
+    change of state of the socket. Between two function calls, an infinite number of segments
+    can be potentially receive. The reception of one or zero segment is modeled by the function
+    \lstinline[language=Ada]{Tcp_Process_One_Segment} in term of change of states. Then we have
+    to consider the iteration of \lstinline[language=Ada]{Tcp_Process_One_Segment} in order to
+    compute the result of the reception of multiple message.
 
-    \subsubsection{Concurrency}
+    \begin{algorithm}[t]
+        \SetKwFunction{ProcessSegm}{\lstinline[language=Ada]{Tcp\_Process\_Segment}}
+        \SetKwProg{Fn}{function}{}{end}
+        \Fn{\ProcessSegm{Socket}}{
+            $S_{last} := \text{\textit{Socket}}$\;
+            $S := S_{last}$\;
+            \For{$i=1$ \KwTo $3$}{
+                $S_{last} :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$) \;
+                $S := S \cup S_{last}$\;
+            }
+            \KwRet{$S$}\;
+        }
+        \caption{Reflexive and transitive closure of \lstinline[language=Ada]{Tcp_Process_One_Segment}}
+        \label{algo:tcpProcessSegment}
+    \end{algorithm}
 
-    Modeling the concurrency and the interleaving semantic has been the hardest work.
-    The held solution is to have "ghost" code that model the different transition that
-    might have happened when the mutex is released.
+    Let $\rightarrow$ be the transition function in the TCP automaton restrited to the transitions
+    that can be performed by the reception of a message and its automatic response mechanism.
+    Then we have $\rightarrow = \text{\lstinline[language=Ada]{Tcp_Process_One_Segment}}$ and
+    we need to consider the reflexive transitive closure $\rightarrow^*$ of $\rightarrow$, with
+    \[\rightarrow^* = \bigcup_{n\in\mathbb{N}} \rightarrow^n\]
+    Now, by examining the TCP automaton in figure \ref{fig:TCPAutomaton}, we see that all state is
+    only reachable by a state at a distance less than 3 (the maximum path is between SYN-SENT and
+    CLOSE-WAIT if we pass by state SYN-RECEIVED and ESTABLISHED) without action by the user. Since
+    we only consider the transitive closure in term of states, we can significantly reduce the number
+    of iteration of $\rightarrow$ to compute $\rightarrow^*$ and finally we get:
+    \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
+    It follows that algorithm \ref{algo:tcpProcessSegment} is enough good to compute the transitive closure of the function
+    \lstinline[language=Ada]{Tcp_Process_One_Segment} by taking advantage of the fact that SPARK can unroll
+    small loops to obtain more precise results.
 
-    The interleaving semantic is also considered for synchronous events, when the code
-    wait for a particular event.
+    In the user functions, we have added a call to \lstinline[language=Ada]{Tcp_Process_Segment} everywhere
+    there is a call to \lstinline{Os_Acquire_Mutex (Net_Mutex)}. Doing that helps to ensure that all the
+    possible that all the possible input states are have been considered.
 
-    If we consider the state transition given by the code that receive the message, we
-    are driven to consider a transition relation, denotes as $\rightarrow$. This relation
-    is reflexive because a received message can be drop if it doesn't contain useful
-    information. Then, for all TCP state $S$, $S \rightarrow S$. As multiples messages
-    can be received, the transitive closure of $\rightarrow$ is considered:
-    $\rightarrow^* = \bigcup_{n\in\mathbb{N}^*} \rightarrow^n$ which can be simplified to:
-    \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
-    since every state is only reachable by a state at a distance at most of 3, which make
-    the calculation of the transition closure easier by SPARK once we have $\rightarrow$ by
-    unrolling loops.
+    \subsubsection{Concurrency: Synchronous exchange}
+
+    The other mechanism to deal with concurrency is probably more difficult to apprehend and the
+    use of the function \lstinline[language=Ada]{Tcp_Process_One_Segment} is even more noticeable.
+
+    The C code contains a function \lstinline[language=C]{tcpWaitForEvents} in the file \texttt{tcp\_misc.c}
+    that checks if the event is true when the function is called, by calling the function
+    \lstinline[language=C]{tcpUpdateEvents}. If the wait is not completed at the time the function is
+    called, then the mutex that was previously locked is released until the event become waited happen.
+    Then, everything can happen meanwhile.
+
+    In the code, the function \lstinline[language=C]{tcpUpdateEvents} is called at different locations
+    and can raise the desired event. In particular this function is called each time the state of
+    the socket is changed. This is sufficient for our purpose since we are only interested in the state
+    changes.
+
+    When a segment is received, the function \lstinline[language=C]{tcpUpdateEvents} is called if a
+    change of state happens. Then, we can consider the algorithm \ref{algo:waitForEvents} to compute the set of possible
+    final states after the wait. \textit{(This is done in the function \lstinline[language=Ada]{Tcp_Wait_For_Events_Proof},
+    the function dedicated to proof in file \texttt{tcp\_misc\_binding.adb}.)}
+
+    \begin{algorithm}[t]
+        \SetKwFunction{ProcessSegm}{\lstinline[language=Ada]{Tcp\_Wait\_For\_Events}}
+        \SetKwProg{Fn}{function}{}{end}
+        \Fn{\ProcessSegm{Socket, Event, Event\_Mask}}{
+            $S_{last} := \text{\textit{Socket}}$\;
+            $S := S_{last}$\;
+            $E :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$)\;
+            \If{$E \wedge \text{Event\_Mask} \neq 0$}{
+                \KwRet{$S$}\;
+            }
+            \For{$i=1$ \KwTo $3$}{
+                $S_{last} :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$) \;
+                $S := S \cup S_{last}$\;
+                $E :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$)\;
+                \If{$E \wedge \text{Event\_Mask} \neq 0$}{
+                    \KwRet{$S$}\;
+                }
+            }
+            \KwRet{$\emptyset$}\;
+        }
+        \caption{Function to compute the possible state after when waiting for a particular event.}
+        \label{algo:waitForEvents}
+    \end{algorithm}
+
+    We can compute precisely the state reach for each event waited thanks to the fact that SPARK
+    unroll loops.
+
+    \textcolor{red}{\paragraph{Weakness} If two threads want to lock the same mutex at the same time, which one win?}
+
+    \section{Future work}
 

### 2020-07-20 klee/established/established.c
Finish the work with klee to check the possible states that can be reached when a segment is received

@@ -1,10 +1,9 @@
 #include "core/socket.h"
 #include "klee/klee.h"
 #include "core/tcp_misc.h"
-#include <assert.h>
 #include <stdlib.h>
+#include <assert.h>
 #include "model.h"
-
 
 void tcpStateEstablished(Socket *socket, TcpHeader *segment,
    const NetBuffer *buffer, size_t offset, size_t length)
@@ -70,65 +69,6 @@ void tcpStateEstablished(Socket *socket, TcpHeader *segment,
    tcpNagleAlgo(socket, flags);
 }
 
-void tcpStateSynReceived(Socket *socket, TcpHeader *segment,
-                         const NetBuffer *buffer, size_t offset, size_t length)
-{
-    //Debug message
-    // TRACE_DEBUG("TCP FSM: SYN-RECEIVED state\r\n");
-
-    //First check sequence number
-    if (tcpCheckSequenceNumber(socket, segment, length))
-        return;
-
-    //Check the RST bit
-    if (segment->flags & TCP_FLAG_RST)
-    {
-        //Return to CLOSED state
-        tcpChangeState(socket, TCP_STATE_CLOSED);
-
-        //Number of times TCP connections have made a direct transition to the
-        //CLOSED state from either the SYN-SENT state or the SYN-RECEIVED state
-        //MIB2_INC_COUNTER32(tcpGroup.tcpAttemptFails, 1);
-        //TCP_MIB_INC_COUNTER32(tcpAttemptFails, 1);
-
-        //Return immediately
-        return;
-    }
-
-    //Check the SYN bit
-    if (tcpCheckSyn(socket, segment, length))
-        return;
-
-    //If the ACK bit is off drop the segment and return
-    if (!(segment->flags & TCP_FLAG_ACK))
-        return;
-
-    //Make sure the acknowledgment number is valid
-    if (segment->ackNum != socket->sndNxt)
-    {
-        //If the segment acknowledgment is not acceptable, form a reset
-        //segment and send it
-        tcpSendSegment(socket, TCP_FLAG_RST, segment->ackNum, 0, 0, FALSE);
-
-        //Drop the segment and return
-        return;
-    }
-
-    //Update the send window before entering ESTABLISHED state (refer to
-    //RFC 1122, section 4.2.2.20)
-    socket->sndWnd = segment->window;
-    socket->sndWl1 = segment->seqNum;
-    socket->sndWl2 = segment->ackNum;
-
-    //Maximum send window it has seen so far on the connection
[…]

### 2020-07-20 klee/last_ack/last_ack.c

@@ -1,0 +1,107 @@
+#include "core/socket.h"
+#include "klee/klee.h"
+#include "core/tcp_misc.h"
+#include <assert.h>
+#include <stdlib.h>
+#include "model.h"
+
+void tcpStateLastAck(Socket *socket, TcpHeader *segment, size_t length)
+{
+   //Debug message
+   // TRACE_DEBUG("TCP FSM: LAST-ACK state\r\n");
+
+   //First check sequence number
+   if(tcpCheckSequenceNumber(socket, segment, length))
+      return;
+
+   //Check the RST bit
+   if(segment->flags & TCP_FLAG_RST)
+   {
+      //Enter CLOSED state
+      tcpChangeState(socket, TCP_STATE_CLOSED);
+      //Return immediately
+      return;
+   }
+
+   //Check the SYN bit
+   if(tcpCheckSyn(socket, segment, length))
+      return;
+   //If the ACK bit is off drop the segment and return
+   if(!(segment->flags & TCP_FLAG_ACK))
+      return;
+
+   //The only thing that can arrive in this state is an
+   //acknowledgment of our FIN
+   if(segment->ackNum == socket->sndNxt)
+   {
+      //Enter CLOSED state
+      tcpChangeState(socket, TCP_STATE_CLOSED);
+   }
+}
+
+int main()
+{
+    // Initialisation
+    socketInit();
+
+    Socket *socket;
+    IpPseudoHeader *pseudoHeader;
+    TcpHeader *segment;
+    size_t length;
+    uint32_t ackNum, seqNum;
+    uint8_t flags, dataOffset;
+    uint16_t checksum, srcPort;
+    struct socketModel *sModel;
+
+    segment = malloc(sizeof(TcpHeader) + sizeof(u_int8_t));
+    pseudoHeader = malloc(sizeof(IpPseudoHeader));
+
+    // creation of a TCP socket
+    socket = socketOpen(SOCKET_TYPE_STREAM, SOCKET_IP_PROTO_TCP);
+
+    // TODO: See if it is the correct way to process.
+    // Maybe open the connection by following the real call of the function
+    // is a better way to do.
+    tcpChangeState(socket, TCP_STATE_LAST_ACK);
+
+    // Creation of a "random" incomming package
+    // We assume the pseudo header is filled with ipv4 addresses
+
+    klee_make_symbolic(&ackNum, sizeof(ackNum), "ackNum");
+    klee_make_symbolic(&seqNum, sizeof(seqNum), "seqNum");
[…]

### 2020-07-20 klee/time_wait/time_wait.c

@@ -1,0 +1,124 @@
+#include "core/socket.h"
+#include "klee/klee.h"
+#include "core/tcp_misc.h"
+#include <assert.h>
+#include <stdlib.h>
+#include "model.h"
+#include "core/tcp_timer.h"
+
+
+void tcpStateTimeWait(Socket *socket, TcpHeader *segment, size_t length)
+{
+   //Debug message
+   // TRACE_DEBUG("TCP FSM: TIME-WAIT state\r\n");
+
+   //First check sequence number
+   if(tcpCheckSequenceNumber(socket, segment, length))
+      return;
+
+   //Check the RST bit
+   if(segment->flags & TCP_FLAG_RST)
+   {
+      //Enter CLOSED state
+      tcpChangeState(socket, TCP_STATE_CLOSED);
+
+      //Dispose the socket if the user does not have the ownership anymore
+      if(!socket->ownedFlag)
+      {
+         //Delete the TCB
+         tcpDeleteControlBlock(socket);
+         //Mark the socket as closed
+         socket->type = SOCKET_TYPE_UNUSED;
+      }
+
+      //Return immediately
+      return;
+   }
+
+   //Check the SYN bit
+   if(tcpCheckSyn(socket, segment, length))
+      return;
+   //If the ACK bit is off drop the segment and return
+   if(!(segment->flags & TCP_FLAG_ACK))
+      return;
+
+   //The only thing that can arrive in this state is a retransmission
+   //of the remote FIN. Acknowledge it and restart the 2 MSL timeout
+   if(segment->flags & TCP_FLAG_FIN)
+   {
+      //Send an acknowledgment for the FIN
+      tcpSendSegment(socket, TCP_FLAG_ACK, socket->sndNxt, socket->rcvNxt, 0, FALSE);
+      //Restart the 2MSL timer
+      tcpTimerStart(&socket->timeWaitTimer, TCP_2MSL_TIMER);
+   }
+}
+
+int main()
+{
+    // Initialisation
+    socketInit();
+
+    Socket *socket;
+    IpPseudoHeader *pseudoHeader;
+    TcpHeader *segment;
+    size_t length;
+    uint32_t ackNum, seqNum;
+    uint8_t flags, dataOffset;
+    uint16_t checksum, srcPort;
+    struct socketModel *sModel;
+
+    segment = malloc(sizeof(TcpHeader) + sizeof(u_int8_t));
+    pseudoHeader = malloc(sizeof(IpPseudoHeader));
+
+    // creation of a TCP socket
[…]

### 2020-07-21 report/report.tex
Update report

@@
 \usetikzlibrary{arrows.meta, shapes.callouts}
+\usetikzlibrary{decorations.pathreplacing}
 
@@
 \usepackage[ruled]{algorithm2e}
+
+\usepackage[backend=biber]{biblatex}
+\addbibresource{biblio.bib}
+
+\DeclareMathOperator{\Pre}{Pre}
+\DeclareMathOperator{\Post}{Post}
 
@@
     protocols, such as SSL/TLS which is built on top of TCP and designed
-    to provide a security layer to TCP that have been the subject of lots of
-    works like miTLS \cite{}, nothing has been done to make a secure TCP
-    implementation. Nonetheless, TLS security can only be ensured if the underlying
-    TCP implementation is free of bugs.
+    to provide a security layer to TCP with for example the work done in
+    miTLS~\cite{bhargavan2013implementing}, or in lower level with the work
+    on RecordFlux to safely parse data segment~\cite{Reiher_2020}, nothing has
+    been done to make a secure TCP implementation. Nonetheless, TLS security
+    can only be ensured if the underlying TCP implementation is free of bugs.
 
@@
     with a dozen of OS. The code of the library is written in C and contains
-    plenty protocols from low-level network layers like IPv4/IPv6, to transport layer
-    and application layer like DHCP or HTTP. In particular an implementation of
+    plenty protocols from low-level network layers like IPv4/IPv6, to transport layers
+    and application layers like DHCP or HTTP. In particular an implementation of
     the TCP protocol is provided.
@@
 
-    The aim of this work was to translate some parts of the CycloneTCP stack
+    The aim of this work was to translate some parts of the CycloneTCP TCP/IP stack
     in Ada/SPARK to improve the safety of the code. In particular we focused
-    on the TCP protocol to ensure that the norm is respected in this work.
+    on the TCP protocol to ensure that the norm is respected in the code.
 
     This report is a synthesis of the challenge of verifying a such protocol
-    and the solution I found to go toward a secure protocol.
+    and the solution we found during the internship to go toward a secure protocol.
 
@@
     As mentionned in the introduction, TCP protocol is the base of a lot of other protocols,
-    in particular SSL/TLS.
+    in particular SSL/TLS. These higher protocols cannot be secure if the underlying TCP protocol
+    has bugs. An incorrect implementation of the TCP protocol can lead to a crash
+    in the communication if two machines cannot communicate together, error in
+    retransmissions, or even security failures. CycloneTCP is designed for embedded
+    code, and a failure in the code can be critical.
 
@@
     information, like for example the local or the remote IP address as well as
-    the state of the connection. In Ada, a socket will be represented by a pointer
-    to a record type:
+    the state of the connection and other information require by the TCP protocol.
+    In Ada, a socket will be represented by a pointer to a record type:
 
@@
 
-    A socket is also the structure that is manipulated by the user as an opaque structure,
-    through an interface to perform operations on it and perform operations on the
+    A socket is also the structure that is manipulated by the users as an opaque structure,
+    through an API to perform operations on the socket itself and on the
     global environment by sending messages in the network for example.
@@
 
-    TCP is defined by the norm RFC 793 \cite{} in a high level language.
+    TCP is defined by the norm RFC 793 \cite{rfc793} in a high level language.
     TCP is a realiable, ordered and error-checked connection oriented protocol.
@@
 
-    As it can be seen in the state machine, there exists different states that are
-    subject to change when a message is segment is sent or received or when an action
-    is done by the user. A label $A/B$ on an edge refers to a stimuli performed on the socket
-    (the receiption of a message, a user action...) for $A$ and a flag sent for $B$.
-    More precisely, for $A$, a text in italic shape refers to an action perform by
-    the user (a call of a user function) or when it's a timeout, by a timer and a ACK,
-    SYN or FIN refers to a flag contained in a received message. The flags in $B$ refer
-    to the flags sent in response of the stimuli $A$.  Without giving
-    unnecessary details on the format of a TCP header, we can describe the different
-    flags that can be contained in a TCP header:
+    As it can be seen in the state machine diagram in figure \ref{fig:TCPAutomaton},
+    a socket can take different states that are
+    subject to change when a message (or a segment) is sent or received or when an action
+    is performed by the user. A label $A/B$ on an edge in the graph refers to a stimuli performed on the socket
+    (the reception of a message, a user action...) for $A$ and a flag sent for $B$.
+    More precisely, for $A$, a text in italic shape refers to an action performed by
+    the user (a call of a user function) or when it's a timeout the action is automatically
+    performed by a timer, and a ACK, SYN or FIN refers to a flag contained in a received message.
+    The flags in $B$ refers to the flags sent in response of the stimuli $A$.
+    Without giving unnecessary details on the format of a TCP header, we can
+    describe some flags that can be contained in a TCP header in order to help the understanding
+    of the figure \ref{fig:TCPAutomaton}:
     \begin{description}
@@
     Different states can be taken by a socket during its lifetime depending on the state
-    of the connection, and also depending on the state of the remote TCP state. Each
-    state has its own signification:
+    of the connection. The state also gives information about the state of the remote TCP.
+    Each state has its own signification:
     \begin{description}
@@
 
-    For more details on how two TCP can establish a connection, the reader can refer
+    For more details on how two TCP can establish or close a connection, the reader can refer
     to the TCP norm where pleinty of scenario are described and explained, and more
@@
     three threads: one for the \emph{user calls}, one for the \emph{arriving segments}
-    and one for the \emph{timers}. This design has been adopted in CycloneTCP.
+    and one for the \emph{timers}. This design has also been adopted in CycloneTCP.
 
@@
     In this threads, the received segments are processed. The corresponding messages are sent back.
-    Transition between state can happen here. For example, this is the case when a message containing
-    a SYN is received when the socket is in the LISTEN state. It will automatically send a SYN + ACK in
-    response, and change its state for SYN-RECEIVED.
+    Transitions between states can happen here when a message is received. For example,
+    this is the case when a message containing a SYN is received when the socket is in
+    the LISTEN state. It will automatically send a SYN + ACK in response, and change
+    its state for SYN-RECEIVED.
 
@@
     The timers control the timeouts like the retransmission timeout to retransmit a message, or
-    the time-wait timeout to close the connection after an amont of time. Then, transition can also
+    the time-wait timeout to close the connection after an amont of time. Then, transitions can also
     happen in this threads.
@@
     To sum up, a single socket can be seen as a shared structure over multiples threads, that can all
-    manipulate it and change its fields. The schema illustrates that.
+    manipulate it and change the value of its fields. The schema in figure \ref{fig:TCPConcurrency} illustrates that.
 
@@
 
+    \begin{figure}
         \begin{tikzpicture}
@@
 
-        \draw [->] (4.3,2) -- (4.9,2);
-        \node [anchor=west] at (5,2) {Operation};
+            \draw[decorate, decoration={brace, amplitude=10pt}] (7.5, 0) -- (7.5, -1.8)
+                node [black,midway,xshift=0.55cm, rotate=-90] {\textit{Threads}};
+
+            \draw [->] (4.3,2.12) -- (4.9,2.12);
+            \node [anchor=west, text width=2.7cm] at (5,2.2) {Operations on the socket};
             \draw[<->, dashed] (4.3,1.3) -- (4.9,1.3);
@@
         \end{tikzpicture}
+        \caption{Concurrency in the TCP protocol.}
+        \label{fig:TCPConcurrency}
+    \end{figure}
 
@@
 
-    Some choices have had to be done in regard the short time devolved to the internship, and
-    not every aspect of the TCP norm has been prooved. However, some aspects have been selected
+    Some choices have had to be done in regard to the short time devolved to the internship, and
+    not every aspect of the TCP norm has been proved yet. However, some aspects have been selected
     for the challenge or the importance they represent. A list of these aspects is presented
@@
     One of the main problems pointed by the main writter of the CycloneTCP library is that users
-    do not correctly used the API supplied. As a result, some calls of
+    do not correctly use the API supplied. As a result, some calls of
     functions can be incorrect, either because the arguments are incorrect, or because the
@@
     do what they are supposed to do. It is a part of the job that has been done. More
-    especially we have wanted to checked if the transition machine is respected by all the
-    user functions. We have also tried to ensure the correctness of the function called
-    regarding to the TCP state of the socket before the call to respect the specifications
-    given in the norm.
+    especially we have wanted to check if the state machine is respected by all the
+    user functions in the transition they do. We have also tried to ensure the correctness
+    of the function called regarding to the TCP state of the socket before the call
+    to respect the specifications given in the norm.
 
@@
     a relation over the oder of which functions have to be called, then we want that the
-    post-condition of $f_1$ implies the precondition of $f_2$.
+    post-condition of $f_1$ implies the precondition of $f_2$, \textit{i.e.} $\Pre(f_2) \subseteq \Post(f_1)$.
 
@@
     \lstinline[language=Ada]{Socket_Send}. The function \lstinline[language=Ada]{Socket_Connect}
-    try to connect to a distant TCP. If the connection succeed, the remote IP address of the distant
+    try to connect to a distant TCP. If the connection succeeds, the remote IP address of the distant
     TCP is set in the field \lstinline[language=Ada]{S_Remote_Ip_Addr} of the \lstinline[language=Ada]{Sock}
@@
     procedure \lstinline[language=Ada]{Socket_Connect}. The post-condition filters the cases
-    where \lstinline{Error} is \lstinline{NO_ERROR} and if \lstinline{Error} takes another value.
+    where \lstinline{Error} is \lstinline{NO_ERROR} and where \lstinline{Error} takes another value.
     As a result, a user who would write an incorrect code such as
@@
 
+    In order to verify the transitions we have read the norm to extract informations about
+    the allowed transitions, and we have added contracts in the function
+    \lstinline[language=Ada]{tcpChangeState}. Since all the transitions are made through
+    this function, an incorrect transition will be warned by SPARK.
+
+    As mentionned, since most of transitions are done outside the user functions, a big part
+    of this work was to consider other transitions in other threads and we present how we have
+    done it in the following sections.
+
     \subsubsection{Overview of the concurrency challenge}
@@
     To ensure the safety of our library, we need to consider all what can happen everywhere in
-    the library. All our functions are protected by mutex, which means that only one operation
+    the library. All the functions are protected by mutex, which means that only one operation
     can be performed at the same time on a socket. Interractions must be considered at two
     locations: between the function calls, when the mutex is released and during the function
-    call, when the program waits for an event. We will explore in the two next sections how we
-    have dealt with this two different concurrent mechanisms.
+    call, when the program waits for an event. We will explore in the sections \ref{section:asynchronous}
+    and \ref{section:synchronous} how we have dealt with this two different concurrent mechanisms.
 
-    Before everything else, let's see where the interference mainly comes from, and how we have
+    Before everything else, let's see where the interferences mainly come from, and how we have
     done to ensure that all have been considered.
@@
     strongly encouraged to have a look at this file. The main function of this file
-    \texttt{tcpProcessSegment} look for the socket corresponding to the received segment, and then
+    \texttt{tcpProcessSegment} looks for the socket corresponding to the received segment, and then
     according to the TCP state of the socket, process the segment as expected by calling one of
     the function \texttt{tcpState<StateName>}. The functions familly \texttt{tcpState<StateName>}
-    check the information contained in the segment and can perform change of state in the socket
+    check the information contained in the segment and can perform a change of state in the socket
     struct depending on the flags received.
 
-    In these functions, change of state can be performed as mentionned, through the function
+    In these functions, a change of state can be performed, as mentioned, thanks to the function
     \texttt{tcpChangeState}. To summarize what is done when a message is received, rather than
     reading the code and locate all the calls to functions \texttt{tcpChangeState}, we have used
-    Klee\footnote{\url{https://klee.github.io}}. It helped to find behavior that had not been
-    imagined at first. Anyway, we cannot state that all the work done with klee is complete, and
+    Klee\footnote{\url{https://klee.github.io}}. It helped to find behaviors that had not been
+    imagined at first. Anyway, we cannot state that all the work done with Klee is complete, and
     we would need to rewrite this part in SPARK to have a code formally proved.
 
-    All what is related to klee is present in the folder \texttt{klee/} and can be compiled and run
+    All what is related to Klee is present in the folder \texttt{klee/} and can be compiled and run
     thanks to the makefile to reproduce the results. Roughtly, what is done is creating a random
@@
 
-    Finally, all the result found by klee has been reported in the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
+    Finally, all the results found by Klee has been reported as a post-condition of the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
     This function is essential for all the verification of concurrent parts and is reused everywhere
-    the concurrency is involved to compute the results.
-
-    % The functions used to process the segments are written in C and we have not changed
-    % them. We have just extracted some information, like the changes of state that can be
-    % produced by the reception of a message.
-
-    % To ensure the correctness of the extracted information, we have used Klee, a symbolic
-    % virtual machine. It allows to consider all the paths taken in the function and check
-    % if the contracts written in SPARK respects the C implementation.
-    % \textbf{\itshape Show the example where I found an error.}
-
-    % A goal to achieve would be to rewrite this part in SPARK to have more guaranties.
-    % However, Klee is a tool that helps to find errors.
+    the concurrency is involved to compute possible interferences. It follows that the safety of the code rely on the
+    confidence we have in this function.
 
     \subsubsection{Concurrency: asynchronous changes of state}
+    \label{section:asynchronous}
 
-    Between the user functions calls, segment can be received and these receptions can drive to
-    change of state of the socket. Between two function calls, an infinite number of segments
+    Between the user functions calls, segments can be received and these receptions can drive to
+    changes of the state of the socket. Between two function calls, an infinite number of segments
     can be potentially receive. The reception of one or zero segment is modeled by the function
@@
     compute the result of the reception of multiple message.
+
+    Let $\rightarrow$ be the transition function in the TCP automaton restrited to the transitions
+    that can be performed by the reception of a message and its automatic response mechanism.
+    Then we have $\rightarrow = \text{\lstinline[language=Ada]{Tcp_Process_One_Segment}}$ and
+    we need to consider the reflexive transitive closure $\rightarrow^*$ of $\rightarrow$, with
+    \[\rightarrow^* = \bigcup_{n\in\mathbb{N}} \rightarrow^n\]
+    Now, by examining the TCP automaton in figure \ref{fig:TCPAutomaton}, we see that all state is
+    only reachable by a state at a distance less than 3 (the maximum path is between SYN-SENT and
+    CLOSE-WAIT if we pass by state SYN-RECEIVED and ESTABLISHED) without user action. Since
+    we only consider the transitive closure in term of states, we can significantly reduce the number
+    of iteration of $\rightarrow$ to compute $\rightarrow^*$ and finally we get:
+    \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
+    It follows that algorithm \ref{algo:tcpProcessSegment} is enough good to compute the transitive closure of the function
+    \lstinline[language=Ada]{Tcp_Process_One_Segment} by taking advantage of the fact that SPARK can unroll
+    small loops to obtain more precise results.
 
@@
 
-    Let $\rightarrow$ be the transition function in the TCP automaton restrited to the transitions
-    that can be performed by the reception of a message and its automatic response mechanism.
-    Then we have $\rightarrow = \text{\lstinline[language=Ada]{Tcp_Process_One_Segment}}$ and
-    we need to consider the reflexive transitive closure $\rightarrow^*$ of $\rightarrow$, with
-    \[\rightarrow^* = \bigcup_{n\in\mathbb{N}} \rightarrow^n\]
-    Now, by examining the TCP automaton in figure \ref{fig:TCPAutomaton}, we see that all state is
-    only reachable by a state at a distance less than 3 (the maximum path is between SYN-SENT and
-    CLOSE-WAIT if we pass by state SYN-RECEIVED and ESTABLISHED) without action by the user. Since
-    we only consider the transitive closure in term of states, we can significantly reduce the number
-    of iteration of $\rightarrow$ to compute $\rightarrow^*$ and finally we get:
-    \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
-    It follows that algorithm \ref{algo:tcpProcessSegment} is enough good to compute the transitive closure of the function
-    \lstinline[language=Ada]{Tcp_Process_One_Segment} by taking advantage of the fact that SPARK can unroll
-    small loops to obtain more precise results.
-
     In the user functions, we have added a call to \lstinline[language=Ada]{Tcp_Process_Segment} everywhere
@@
     \subsubsection{Concurrency: Synchronous exchange}
+    \label{section:synchronous}
 
@@
     \lstinline[language=C]{tcpUpdateEvents}. If the wait is not completed at the time the function is
-    called, then the mutex that was previously locked is released until the event become waited happen.
+    called, then the mutex that was previously locked is released until the waited event happens.
     Then, everything can happen meanwhile.
@@
     When a segment is received, the function \lstinline[language=C]{tcpUpdateEvents} is called if a
-    change of state happens. Then, we can consider the algorithm \ref{algo:waitForEvents} to compute the set of possible
+    change of state happens. Then, we can consider the algorithm \ref{algo:waitForEvents} that reuse
+    \lstinline[language=Ada]{Tcp_Process_One_Segment} to compute the most precise set of possible
     final states after the wait. \textit{(This is done in the function \lstinline[language=Ada]{Tcp_Wait_For_Events_Proof},
@@
 
-    We can compute precisely the state reach for each event waited thanks to the fact that SPARK
-    unroll loops.
+    We can compute precisely the states reach for each event waited thanks to the fact that SPARK
+    unrolls loops.
 
-    \textcolor{red}{\paragraph{Weakness} If two threads want to lock the same mutex at the same time, which one win?}
+    \textcolor{red}{\paragraph{Weakness} If two threads want to lock the same mutex at the same time, which one win?
+    This is probably something to discuss with Clément, because the function \texttt{tcpWaitForEvents} could have a
+    different semantic than what we imagine at first.}
 
@@
 
+    Other improvements that could be done in the future:
+    \begin{itemize}
+        \item Fully translate the file \texttt{tcp\_misc.c} in Ada to have more guarantees.
+        \item Same for the file \texttt{tcp\_fsm.c} that has been investigate with klee. But
+        it is not enough and we need more guarantees for this file, in order to prove the
+        interactions between threads first, and also to prove the correctness of this part
+        of code w.r.t. the TCP norm.
+    \end{itemize}
+
+    \printbibliography
+
 \end{document}

### 2020-07-21 report/report.tex
Typo/English review

@@
     Ensuring the safety of the TCP/IP stack is essential for the safety of
-    a lot of machines. While a lot of work has been done on higher level
+    a lot of machines. While a lot of work has been done on formally verifying higher level
     protocols, such as SSL/TLS which is built on top of TCP and designed
     to provide a security layer to TCP with for example the work done in
-    miTLS~\cite{bhargavan2013implementing}, or in lower level with the work
-    on RecordFlux to safely parse data segment~\cite{Reiher_2020}, nothing has
-    been done to make a secure TCP implementation. Nonetheless, TLS security
+    miTLS~\cite{bhargavan2013implementing}, or at a lower level with the work
+    on RecordFlux to safely parse data segments~\cite{Reiher_2020}, nothing has
+    been done to formally verify a secure TCP implementation. However, TLS security
     can only be ensured if the underlying TCP implementation is free of bugs.
 
-    CycloneTCP is a library developped by Oryx Embedded for embedded platforms.
+    CycloneTCP is a library developped by the company Oryx Embedded for embedded platforms.
     A large number of platforms are supported, and the library can be used
-    with a dozen of OS. The code of the library is written in C and contains
-    plenty protocols from low-level network layers like IPv4/IPv6, to transport layers
+    with a dozen of OS. The code of the library is written in C and implements
+    many protocols from low-level network layers like IPv4/IPv6, to transport layers
     and application layers like DHCP or HTTP. In particular an implementation of
@@
     SPARK 2014 is a programming language designed as a subset of Ada that helps in
-    making high reliability sofwares by providing powerful static analysis methods.
+    making high reliability sofware by providing powerful static analysis methods.
     SPARK is able to detect uninitialized variables with control flow analysis,
     and can ensure the absence of run-time errors,
-    but also, based on SMT-solver, functionnal behavior can be specified for every
-    functions and mathematically proved.
+    but also, based on SMT-solver, functional behavior can be specified for every
+    function and mathematically proved.
 
@@
 
-    This report is a synthesis of the challenge of verifying a such protocol
-    and the solution we found during the internship to go toward a secure protocol.
+    This report is a synthesis of the challenge of verifying such a protocol
+    and the solution we found during the internship to go towards a secure protocol.
 
@@
 
-    \subsection{Why verifying a TCP stack?}
+    \subsection{Why verify a TCP stack?}
 
-    As mentionned in the introduction, TCP protocol is the base of a lot of other protocols,
+    As mentioned in the introduction, TCP protocol is the base of a lot of other protocols,
     in particular SSL/TLS. These higher protocols cannot be secure if the underlying TCP protocol
     has bugs. An incorrect implementation of the TCP protocol can lead to a crash
-    in the communication if two machines cannot communicate together, error in
-    retransmissions, or even security failures. CycloneTCP is designed for embedded
+    in the communication if two machines cannot communicate together, to an error in
+    retransmissions, or even to security failures. CycloneTCP is designed for embedded
     code, and a failure in the code can be critical.
@@
 
-    A socket is nothing more than a data structures that contains the connection
+    A socket is nothing more than a data structure that contains the connection
     information, like for example the local or the remote IP address as well as
-    the state of the connection and other information require by the TCP protocol.
+    the state of the connection and other information required by the TCP protocol.
     In Ada, a socket will be represented by a pointer to a record type:
@@
     TCP is defined by the norm RFC 793 \cite{rfc793} in a high level language.
-    TCP is a realiable, ordered and error-checked connection oriented protocol.
+    TCP is a reliable, ordered and error-checked connection oriented protocol.
     It means that two computers have to open a connection first before sending data.
@@
     subject to change when a message (or a segment) is sent or received or when an action
-    is performed by the user. A label $A/B$ on an edge in the graph refers to a stimuli performed on the socket
+    is performed by the user. A label $A/B$ on an edge in the graph refers to a stimulus performed on the socket
     (the reception of a message, a user action...) for $A$ and a flag sent for $B$.
@@
     performed by a timer, and a ACK, SYN or FIN refers to a flag contained in a received message.
-    The flags in $B$ refers to the flags sent in response of the stimuli $A$.
+    The flags in $B$ refer to the flags sent in response of the stimulus by $A$.
     Without giving unnecessary details on the format of a TCP header, we can
     describe some flags that can be contained in a TCP header in order to help the understanding
-    of the figure \ref{fig:TCPAutomaton}:
+    of figure \ref{fig:TCPAutomaton}:
     \begin{description}
         \item[ACK] Acknowledgment field significant. The last message received by the sender
-        is acknowledge. A field of the header contains the number of this segment.
+        is acknowledged. A field of the header contains the number of this segment.
         \item[SYN] Synchronize sequence number. This flag is sent in order to establish
@@
     For more details on how two TCP can establish or close a connection, the reader can refer
-    to the TCP norm where pleinty of scenario are described and explained, and more
+    to the TCP norm where many scenarios are described and explained, and more
     details are given on the state and segments that can be sent.
@@
     \paragraph{User calls}
-    User calls refers to functions that can be called by the user to control the connection,
-    send, or received data: OPEN, CLOSE, ABORT, SEND, RECEIVED. Transition between state of
-    the connection can happen during the call of these function since there are made to control
+    User calls refer to functions that can be called by the user to control the connection,
+    send, or received data: OPEN, CLOSE, ABORT, SEND, RECEIVED. Transition between states of
+    the connection can happen during the call of these functions since they are intended to control
     the connection.
@@
     \paragraph{Arriving segments}
-    In this threads, the received segments are processed. The corresponding messages are sent back.
+    In this thread, the received segments are processed. The corresponding messages are sent back.
     Transitions between states can happen here when a message is received. For example,
@@
     The timers control the timeouts like the retransmission timeout to retransmit a message, or
-    the time-wait timeout to close the connection after an amont of time. Then, transitions can also
-    happen in this threads.
+    the time-wait timeout to close the connection after an amount of time. Then, transitions can also
+    happen in this thread.
 
-    To sum up, a single socket can be seen as a shared structure over multiples threads, that can all
+    To sum up, a single socket can be seen as a shared structure over multiple threads, that can all
     manipulate it and change the value of its fields. The schema in figure \ref{fig:TCPConcurrency} illustrates that.
@@
             \node [draw, text width=2.5cm, anchor=north] (T2) at (2,0) {\textbf{USER TASK} \\
-                                                    Controling the connection on the user side};
+                                                    Controlling the connection on the user side};
             \node [draw, text width=2.5cm, anchor=north] (T3) at (6, 0) {\textbf{RECEIVED TASK} \\
@@
         messages received are correctly processed in regard of the flags they contain.
-        \item Functionnal correctness of the user functions in regard to their specifications.
+        \item Functional correctness of the user functions in regard to their specifications.
     \end{itemize}
@@
     here and the choices will be motivated in the next subsections. First of all, we have wanted
-    to improve the interface of the functions. The other point was to improved the safety of the
+    to improve the interface of the functions. The other point was to improve the safety of the
     protocol itself.
 
-    \subsubsection{Improve functions interface}
+    \subsubsection{Improve the functional interface}
 
-    One of the main problems pointed by the main writter of the CycloneTCP library is that users
-    do not correctly use the API supplied. As a result, some calls of
-    functions can be incorrect, either because the arguments are incorrect, or because the
-    return code of the previous functions has not been checked, which drive to an
+    One of the main problems pointed to by the main author of the CycloneTCP library is that users
+    do not correctly use the API supplied. As a result, some function
+    calls can be incorrect, either because the arguments are incorrect, or because the
+    return code of the previous calls has not been checked, which lead to an
     incorrect code.
 
-    \subsubsection{Respect to the protocol}
+    \subsubsection{Respect of the protocol}
 
@@
     user functions in the transition they do. We have also tried to ensure the correctness
-    of the function called regarding to the TCP state of the socket before the call
+    of the functions called regarding the TCP state of the socket before the call
     to respect the specifications given in the norm.
@@
     between the multiples threads: synchronous and asynchronous.
-    All these interraction have to be considered if we want to write correct contracts
+    All these interactions have to be considered if we want to write correct contracts
     in regard of what could be done in other threads.
 
-    As SPARK does not have a native mode to deal with concurrency, we have to
+    As SPARK does not have a native mode to deal precisely with interactions related to concurrency, we have to
     model these different interactions by hands by writing assertions that
@@
     the problem has been encountered when we tried to write contracts for the correctness
-    of the user functions. More details and explainations will be given in the following
+    of the user functions. More details and explanations will be given in the following
     section, where we will investigate the solution found to these technical problems.
@@
 
-    This part mainly focussed on the high level user functions, the socket functions.
-    The socket functions are the one manipulated by the user to perform operations on
+    This part mainly focusses on the high level user functions, the socket functions.
+    The socket functions are the one called by the user to perform operations on
     the network. They are located in the files \texttt{socket\_interface.ad(b|s)}.
 
-    It is clear that there exist an order on which the functions have to be called.
+    It is clear that there exists an order in which the functions have to be called.
     This order can be determined by the norm, and must respect the order given by the
@@
     If two functions $f_1$ and $f_2$ are ordered such that $f_1 \preceq f_2$ where $\preceq$ is
-    a relation over the oder of which functions have to be called, then we want that the
+    a relation over the order in which functions have to be called, then we want that the
     post-condition of $f_1$ implies the precondition of $f_2$, \textit{i.e.} $\Pre(f_2) \subseteq \Post(f_1)$.
@@
     \lstinline[language=Ada]{Socket_Send}. The function \lstinline[language=Ada]{Socket_Connect}
-    try to connect to a distant TCP. If the connection succeeds, the remote IP address of the distant
+    tries to connect to a distant TCP. If the connection succeeds, the remote IP address of the distant
     TCP is set in the field \lstinline[language=Ada]{S_Remote_Ip_Addr} of the \lstinline[language=Ada]{Sock}
-    structure. When a user call the function \lstinline[language=Ada]{Socket_Send}, we want to ensure that
+    structure. When a user calls the function \lstinline[language=Ada]{Socket_Send}, we want to ensure that
     the connection has already been established. This is why a precondition of this function is
-    \lstinline[language=Ada]{Is_Initialized_Ip(Sock.S_Remote_Ip_Addr)}.
+    \lstinline[language=Ada]{Is_Initialized_Ip(Sock.S_Remote_Ip_Addr)}. This precondition will only
+    be proved if \lstinline[language=Ada]{Socket_Connect} has been previously called.
 
@@
         \end{lstlisting}
-        \caption{An example of how functions calls can be ordered by Pre- and Post-conditions.}
+        \caption{An example of how function calls can be ordered by Pre- and Post-conditions.}
         \label{code:FunOrdered}
@@
     An observation made by Clément Zeller, the main programmer of the library is that customers
-    do not always think to check the return code of the socket user functions, that can fail.
+    do not always think to check the return code of the socket user functions, to deal with the
+    possibility that the call fails.
     If the return code is not checked, some assumption on the result cannot be done.
@@
     The post-conditions as we have written them in SPARK, ensure that the return code is checked
-    before continuating processing. Figure \ref{code:FunOrdered} shows this mechanism for the
-    procedure \lstinline[language=Ada]{Socket_Connect}. The post-condition filters the cases
+    before processing continues. Figure \ref{code:FunOrdered} shows this mechanism for the
+    procedure \lstinline[language=Ada]{Socket_Connect}. The post-condition distinguishes the cases
     where \lstinline{Error} is \lstinline{NO_ERROR} and where \lstinline{Error} takes another value.
@@
 
+    So in the end, the precondition of \lstinline[language=Ada]{Socket_Send} will only
+    be proved if \lstinline[language=Ada]{Socket_Connect} has been previously called, and the code
+    checks after that call that \lstinline[language=Ada]{Error = NO_ERROR}, and the socket
+    \lstinline[language=Ada]{Sock} has not been modified.
+
     % \subsection{Liberties taken with the norm}
@@
 
-    The aim of this part is to explain how we have done to check that the transitions done
-    in a function respect the TCP automaton. In particular, we were interested in verifying
+    The aim of this part is to explain how we check that the transitions done
+    in a function respect the TCP automaton. In particular, we are interested in verifying
     the high level user functions. These functions are not the one in which most of the
-    transitions are done. Even, the verification of these functions is important in order
+    transitions are done. Still, the verification of these functions is important in order
     to guarantee the safety of the whole library. The functions of interest are located in
@@
     \lstinline[language=Ada]{tcpChangeState}. Since all the transitions are made through
-    this function, an incorrect transition will be warned by SPARK.
+    this function, an incorrect transition will lead to a message by SPARK.
 
-    As mentionned, since most of transitions are done outside the user functions, a big part
+    As mentioned, since most transitions are done outside the user functions, a big part
     of this work was to consider other transitions in other threads and we present how we have
@@
     the library. All the functions are protected by mutex, which means that only one operation
-    can be performed at the same time on a socket. Interractions must be considered at two
+    can be performed at the same time on a socket. Interactions must be considered at two
     locations: between the function calls, when the mutex is released and during the function
     call, when the program waits for an event. We will explore in the sections \ref{section:asynchronous}
-    and \ref{section:synchronous} how we have dealt with this two different concurrent mechanisms.
+    and \ref{section:synchronous} how we have dealt with these two different concurrent mechanisms.
 
     Before everything else, let's see where the interferences mainly come from, and how we have
-    done to ensure that all have been considered.
+    ensured that all have been considered.
 
-    \subsubsection{Process segments functions}
+    \subsubsection{Functions that process segments}
 
@@
     \texttt{tcpProcessSegment} looks for the socket corresponding to the received segment, and then
-    according to the TCP state of the socket, process the segment as expected by calling one of
-    the function \texttt{tcpState<StateName>}. The functions familly \texttt{tcpState<StateName>}
+    according to the TCP state of the socket, processes the segment as expected by calling one of
+    the function \texttt{tcpState<StateName>}. The family of functions \texttt{tcpState<StateName>}
     check the information contained in the segment and can perform a change of state in the socket
@@
     reading the code and locate all the calls to functions \texttt{tcpChangeState}, we have used
-    Klee\footnote{\url{https://klee.github.io}}. It helped to find behaviors that had not been
-    imagined at first. Anyway, we cannot state that all the work done with Klee is complete, and
+    KLEE\footnote{\url{https://klee.github.io}}. It helped to find behaviors that we had not 
+    imagined at first. Anyway, we cannot state that all the work done with KLEE is complete, and
     we would need to rewrite this part in SPARK to have a code formally proved.
 
-    All what is related to Klee is present in the folder \texttt{klee/} and can be compiled and run
+    All the work related to KLEE is present in the folder \texttt{klee/} and can be compiled and run
     thanks to the makefile to reproduce the results. Roughtly, what is done is creating a random
-    incomming message, put the socket in the desired state and call the desired function to see what
+    incoming message, put the socket in the desired state and call the desired function to see what
     final states can be obtain.
 
-    Finally, all the results found by Klee has been reported as a post-condition of the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
+    Finally, all the results found by KLEE have been reported as a post-condition of the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
     This function is essential for all the verification of concurrent parts and is reused everywhere
@@
 
-    Between the user functions calls, segments can be received and these receptions can drive to
+    Between the user function calls, segments can be received and these receptions can lead to
     changes of the state of the socket. Between two function calls, an infinite number of segments
-    can be potentially receive. The reception of one or zero segment is modeled by the function
-    \lstinline[language=Ada]{Tcp_Process_One_Segment} in term of change of states. Then we have
+    can be potentially received. The reception of one or zero segments is modeled by the function
+    \lstinline[language=Ada]{Tcp_Process_One_Segment} in terms of change of states. Then we have
     to consider the iteration of \lstinline[language=Ada]{Tcp_Process_One_Segment} in order to
-    compute the result of the reception of multiple message.
+    compute the result of the reception of multiple messages.
 
-    Let $\rightarrow$ be the transition function in the TCP automaton restrited to the transitions
+    Let $\rightarrow$ be the transition function in the TCP automaton restricted to the transitions
     that can be performed by the reception of a message and its automatic response mechanism.
@@
     \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
-    It follows that algorithm \ref{algo:tcpProcessSegment} is enough good to compute the transitive closure of the function
+    It follows that algorithm \ref{algo:tcpProcessSegment} is good enough to compute the transitive closure of the function
     \lstinline[language=Ada]{Tcp_Process_One_Segment} by taking advantage of the fact that SPARK can unroll
@@
     In the user functions, we have added a call to \lstinline[language=Ada]{Tcp_Process_Segment} everywhere
-    there is a call to \lstinline{Os_Acquire_Mutex (Net_Mutex)}. Doing that helps to ensure that all the
-    possible that all the possible input states are have been considered.
+    there is a call to \lstinline{Os_Acquire_Mutex (Net_Mutex)}. Doing that helps to ensure that all 
+    possible input states have been considered.
 
-    \subsubsection{Concurrency: Synchronous exchange}
+    \subsubsection{Concurrency: synchronous exchange}
     \label{section:synchronous}
 
-    The other mechanism to deal with concurrency is probably more difficult to apprehend and the
+    The other mechanism to deal with concurrency is probably more difficult to comprehend and the
     use of the function \lstinline[language=Ada]{Tcp_Process_One_Segment} is even more noticeable.
@@
     \lstinline[language=C]{tcpUpdateEvents}. If the wait is not completed at the time the function is
-    called, then the mutex that was previously locked is released until the waited event happens.
+    called, then the mutex that was previously locked is released until the expected event happens.
     Then, everything can happen meanwhile.
@@
     When a segment is received, the function \lstinline[language=C]{tcpUpdateEvents} is called if a
-    change of state happens. Then, we can consider the algorithm \ref{algo:waitForEvents} that reuse
+    change of state happens. Then, we can consider the algorithm \ref{algo:waitForEvents} that reuses
     \lstinline[language=Ada]{Tcp_Process_One_Segment} to compute the most precise set of possible
@@
 
-    We can compute precisely the states reach for each event waited thanks to the fact that SPARK
+    We can compute precisely the states reached for each expected event thanks to the fact that SPARK
     unrolls loops.

### 2020-07-24 report/report.tex
Update report

@@
 
+\let\state\textsf
+\newcommand\ESTABLISHED{\state{ESTABLISHED}}
+\newcommand\SYNRECEIVED{\state{SYN-RECEIVED}}
+\newcommand\CLOSED{\state{CLOSED}}
+\newcommand\CLOSEWAIT{\state{CLOSE-WAIT}}
+\newcommand\LISTEN{\state{LISTEN}}
+\newcommand\SYNSENT{\state{SYN-SENT}}
+\newcommand\FINWAITONE{\state{FIN-WAIT-1}}
+\newcommand\FINWAITTWO{\state{FIN-WAIT-2}}
+\newcommand\CLOSING{\state{CLOSING}}
+\newcommand\LASTACK{\state{LAST-ACK}}
+\newcommand\TIMEWAIT{\state{TIME-WAIT}}
+
 \begin{document}
@@
 
-    CycloneTCP is a library developped by the company Oryx Embedded for embedded platforms.
+    CycloneTCP is a library developed by the company Oryx Embedded for embedded platforms.
     A large number of platforms are supported, and the library can be used
@@
     SPARK 2014 is a programming language designed as a subset of Ada that helps in
-    making high reliability sofware by providing powerful static analysis methods.
+    making high reliability software by providing powerful static analysis methods.
     SPARK is able to detect uninitialized variables with control flow analysis,
@@
         \node[state, name=synSent, below of=listen, right of=listen, xshift=8em] {SYN\_SENT};
-        \node[state, name=synRcvd, below of=listen, left of=listen, xshift=-8em] {SYN\_RCVD};
+        \node[state, name=synRcvd, below of=listen, left of=listen, xshift=-8em] {SYN\_RECEIVED};
         \node[state, name=established, below of=listen, node distance=14em] {ESTABLISHED};
@@
         \draw (synSent.west) 
-        edge[stateEdge] node[edgeLabel, yshift=1em]{SYN/SYN + ACK} 
+        edge[stateEdge] node[edgeLabel, yshift=1em]{SYN/ACK} 
         (synRcvd.east);
@@
         \draw ($(synRcvd.east) + (0,-1em)$) 
-        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1em, yshift=-1em]{ACK} 
+        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1em, yshift=-1em]{ACK/-} 
         ($(established.north) + (-1em,0)$);
@@
         \draw (finWait1.south) 
-        edge[stateEdge] node[edgeLabel, xshift=-2em]{ACK} 
+        edge[stateEdge] node[edgeLabel, xshift=-2em]{ACK/-} 
         (finWait2.north);
         \draw ($(finWait1.east) + (0,-.5em)$) 
-        edge[stateEdge, bend left=22.5] node[edgeLabel, yshift=1em]{ACK} 
+        edge[stateEdge, bend left=22.5] node[edgeLabel, yshift=1em]{FIN/ACK} 
         (closing.north);
@@
         \draw (closing) 
-        edge[stateEdge] node[edgeLabel, xshift=-2em]{ACK} 
+        edge[stateEdge] node[edgeLabel, xshift=-2em]{ACK/-} 
         (timeWait);
@@
     \begin{description}
-        \item[CLOSED] represents no connection state at all.
-        \item[LISTEN] represents waiting for a connection request from any remote
+        \item[\CLOSED] represents no connection state at all.
+        \item[\LISTEN] represents waiting for a connection request from any remote
         TCP and port.
-        \item[SYN-SENT] represents waiting for a matching connection request
+        \item[\SYNSENT] represents waiting for a matching connection request
         after having sent a connection request.
-        \item[SYN-RECEIVED] represents waiting for a confirming connection
+        \item[\SYNRECEIVED] represents waiting for a confirming connection
         request acknowledgment after having both received and sent a
         connection request.
-        \item[ESTABLISHED] represents an open connection, data received can be
+        \item[\ESTABLISHED] represents an open connection, data received can be
         delivered to the user. The normal state for the data transfer phase
         of the connection.
-        \item[FIN-WAIT-1] represents waiting for a connection termination request
+        \item[\FINWAITONE{}] represents waiting for a connection termination request
         from the remote TCP, or an acknowledgment of the connection
         termination request previously sent.
-        \item[FIN-WAIT-2] represents waiting for a connection termination request
+        \item[\FINWAITTWO{}] represents waiting for a connection termination request
         from the remote TCP.
-        \item[CLOSE-WAIT] represents waiting for a connection termination request
+        \item[\CLOSEWAIT] represents waiting for a connection termination request
         from the local user.
-        \item[CLOSING] represents waiting for a connection termination request
+        \item[\CLOSING{}] represents waiting for a connection termination request
         acknowledgment from the remote TCP.
-        \item[LAST-ACK] represents waiting for an acknowledgment of the
+        \item[\LASTACK{}] represents waiting for an acknowledgment of the
         connection termination request previously sent to the remote TCP
@@
         request).
-        \item[TIME-WAIT] represents waiting for enough time to pass to be sure
+        \item[\TIMEWAIT{}] represents waiting for enough time to pass to be sure
         the remote TCP received the acknowledgment of its connection
@@
 
-    For more details on how two TCP can establish or close a connection, the reader can refer
+    \paragraph{Example of a possible way to establish a connection}
+
+    We present a scenario on how a connection can be established between two TCPs,
+    when one TCP (TCP A) is in the state \CLOSED{} and the other (TCP B) is in the state \LISTEN.
+    \begin{enumerate}
+        \item TCP A wants to initiate a connection with TCP B. It sends a SYN segment to
+            TCP B and passes to the state \SYNSENT{}.
+
+        \item TCP B receives a segment SYN. It has to respond to this segment by a segment
+            containing the flags SYN and ACK (acknowledgment of the received segment).
+            It changes its state for \SYNRECEIVED.
+
+        \item When TCP A receives the segment sent by TCP B containing the flags SYN and ACK,
+            it only has to send back an ACK of this segment to pass in the state \ESTABLISHED.
+
+        \item TCP B receives the ACK of its segment and pass in state \ESTABLISHED{} too.
+    \end{enumerate}
+    At the end of this procedure, the two TCPs are in state \ESTABLISHED{} and can start to
+    send data. Figure \ref{fig:handshake} illustrates the procedure described above. This procedure to open a connection is known as the ``three-way
+    handshake'' in the TCP norm.
+
+    \begin{figure}
+        \centering
+        \begin{tikzpicture}
+            \node (A) at (0, 5) {\bfseries TCP A};
+            \node (B) at (5, 5) {\bfseries TCP B};
+            \draw[dashed] (-.7,4.8) -- (5.7,4.8);
+            \node (TCPA1) at (0, 4.4) {\CLOSED{}};
+            \node (TCPB1) at (5, 3.9) {\LISTEN{}};
+            \node (TCPA2) at (0, 3.4) {\SYNSENT};
+            \node (TCPB2) at (5, 2.9) {\SYNRECEIVED{}};
+            \node (TCPA3) at (0, 2.4) {\ESTABLISHED{}};
+            \node (TCPB3) at (5, 1.9) {\ESTABLISHED{}};
+
+            \draw[->] (TCPA1) -- (TCPA2);
+            \draw[->] (TCPB1) -- (TCPB2);
+            \draw[->] (TCPA2) -- (TCPA3);
+            \draw[->] (TCPB2) -- (TCPB3);
+            \draw[->, dashed] (0.2, 3.9) -- node [midway, fill=white] {\footnotesize SYN} (TCPB1);
+            \draw[->, dashed] (4.8, 3.4) -- node [midway, fill=white] {\footnotesize SYN + ACK} (TCPA2);
+            \draw[->, dashed] (0.2, 2.9) -- node [midway, fill=white] {\footnotesize ACK} (TCPB2);
+        \end{tikzpicture}
+        \par
+        \caption{Three-way hanshake procedure. The solid lines represent the transition between
+        the states and the dashed lines represent the messages sent.}
+        \label{fig:handshake}
+    \end{figure}
+
+    \bigskip
+
+    We want to point out to the reader the differences that exist between the state machine
+    of figure \ref{fig:TCPAutomaton} and the one that can be found in~\cite[23]{rfc793}. In particular
+    a transition between the states \FINWAITONE{} and \TIMEWAIT{} has been added. It reflects the case where
+    the ACK and FIN flags are received in the same segment. It is an allowed behavior by the norm, and
+    also it is closer to what is done in CycloneTCP library.
+
+    For more details on how two TCPs can establish or close a connection, the reader can refer
     to the TCP norm where many scenarios are described and explained, and more
@@
 
-    \subsubsection{Multithreading in the implementation}
+    \subsubsection{Multitasking in the implementation}
 
-    Different threads can interact with a socket. The TCP norm gives an example with
-    three threads: one for the \emph{user calls}, one for the \emph{arriving segments}
+    Different tasks can interact with a socket. The TCP norm gives an example with
+    three tasks: one for the \emph{user calls}, one for the \emph{arriving segments}
     and one for the \emph{timers}. This design has also been adopted in CycloneTCP.
@@
     User calls refer to functions that can be called by the user to control the connection,
-    send, or received data: OPEN, CLOSE, ABORT, SEND, RECEIVED. Transition between states of
+    send, or receive data: OPEN, CLOSE, ABORT, SEND, RECEIVED. Transition between states of
     the connection can happen during the call of these functions since they are intended to control
@@
     \paragraph{Arriving segments}
-    In this thread, the received segments are processed. The corresponding messages are sent back.
+    In this task, the received segments are processed. The corresponding messages are sent back.
     Transitions between states can happen here when a message is received. For example,
     this is the case when a message containing a SYN is received when the socket is in
-    the LISTEN state. It will automatically send a SYN + ACK in response, and change
-    its state for SYN-RECEIVED.
+    the \LISTEN{} state. It will automatically send a SYN + ACK in response, and change
+    its state for \SYNRECEIVED.
 
@@
     the time-wait timeout to close the connection after an amount of time. Then, transitions can also
-    happen in this thread.
+    happen in this task.
 
-    To sum up, a single socket can be seen as a shared structure over multiple threads, that can all
+    To sum up, a single socket can be seen as a shared structure over multiple tasks, that can all
     manipulate it and change the value of its fields. The schema in figure \ref{fig:TCPConcurrency} illustrates that.
 
-    Only one thread can perform an operation on the socket at the same time. The access to the fields of
+    Only one task can perform an operation on the socket at the same time. The access to the fields of
     the socket are protected by mutex.
 
-    Two threads can communicate synchronously or asynchronously together. The synchronous communications are based on
+    Two tasks can communicate synchronously or asynchronously together. The synchronous communications are based on
     the interface provided by the OS, in particular with events.
@@
             \draw[decorate, decoration={brace, amplitude=10pt}] (7.5, 0) -- (7.5, -1.8)
-                node [black,midway,xshift=0.55cm, rotate=-90] {\textit{Threads}};
+                node [black,midway,xshift=0.55cm, rotate=-90] {\textit{Tasks}};
 
@@
         \item Integrity of the messages sent, \textit{i.e.} we could check that the sent
-        message contains the correct flag and the correct sequence and acknowledgement numbers.
+        message contains the correct flag and the correct sequence and acknowledgment numbers.
         \item Integrity of the messages received, \textit{i.e.} we could check that the
@@
 
-    Some choices have had to be done in regard to the short time devolved to the internship, and
+    Some choices have had to be made in regard to the short time devolved to the internship, and
     not every aspect of the TCP norm has been proved yet. However, some aspects have been selected
@@
     calls can be incorrect, either because the arguments are incorrect, or because the
-    return code of the previous calls has not been checked, which lead to an
+    return code of the previous calls has not been checked, which leads to an
     incorrect code.
@@
 
-    The multiples tasks and the changes that can happen at every moment in one
-    task or another make the verification hard. Multiples interactions exist
-    between the multiples threads: synchronous and asynchronous.
+    The multiple tasks and the changes that can happen at every moment in one
+    task or another make the verification hard. Multiple interactions exist
+    between the multiple tasks: synchronous and asynchronous.
     All these interactions have to be considered if we want to write correct contracts
-    in regard of what could be done in other threads.
+    in regard to what could be done in other tasks.
 
@@
     model these different interactions by hands by writing assertions that
-    modify the state of the socket as another thread could do it. In particular
+    modify the state of the socket as another task could do it. In particular
     the problem has been encountered when we tried to write contracts for the correctness
@@
 
-    This part mainly focusses on the high level user functions, the socket functions.
+    This part mainly focuses on the high level user functions, the socket functions.
     The socket functions are the one called by the user to perform operations on
@@
 
-    In order to verify the transitions we have read the norm to extract informations about
+    In order to verify the transitions we have read the norm to extract information about
     the allowed transitions, and we have added contracts in the function
@@
     As mentioned, since most transitions are done outside the user functions, a big part
-    of this work was to consider other transitions in other threads and we present how we have
+    of this work was to consider other transitions in other tasks and we present how we have
     done it in the following sections.
@@
     Almost all the transitions are done in the file \texttt{tcp\_fsm.c}. This file has not been
-    translated in SPARK by a lack of time. Then we do not have strong guaranties on its functions
+    translated in SPARK by a lack of time. Then we do not have strong guarantees on its functions
     behavior. However, for formal verification we need to know what is done in the functions of
@@
     All the work related to KLEE is present in the folder \texttt{klee/} and can be compiled and run
-    thanks to the makefile to reproduce the results. Roughtly, what is done is creating a random
+    thanks to the makefile to reproduce the results. Roughly, what is done is creating a random
     incoming message, put the socket in the desired state and call the desired function to see what
-    final states can be obtain.
+    final states can be obtained.
 
@@
     This function is essential for all the verification of concurrent parts and is reused everywhere
-    the concurrency is involved to compute possible interferences. It follows that the safety of the code rely on the
-    confidence we have in this function.
+    the concurrency is involved to compute possible interferences. It follows that the safety of the code
+    relies on the confidence we have in this function.
 
@@
     Now, by examining the TCP automaton in figure \ref{fig:TCPAutomaton}, we see that all state is
-    only reachable by a state at a distance less than 3 (the maximum path is between SYN-SENT and
-    CLOSE-WAIT if we pass by state SYN-RECEIVED and ESTABLISHED) without user action. Since
+    only reachable by a state at a distance less than 3 (the maximum path is between \SYNSENT{} and
+    \CLOSEWAIT{} if we pass by the states \SYNRECEIVED{} and \ESTABLISHED{}) without user action. Since
     we only consider the transitive closure in term of states, we can significantly reduce the number
@@
     \lstinline[language=Ada]{Tcp_Process_One_Segment} by taking advantage of the fact that SPARK can unroll
-    small loops to obtain more precise results.
+    small loops to obtain more precise results. The contract of \lstinline[language=Ada]{Tcp_Process_Segment} is
+    manually written and proved with SPARK.
 
@@
 
-    In the code, the function \lstinline[language=C]{tcpUpdateEvents} is called at different locations
-    and can raise the desired event. In particular this function is called each time the state of
+    In the code, the function \lstinline[language=C]{tcpUpdateEvents} is called at different locations; it
+    updates the events true for the socket and it can raise the desired event.
+    In particular this function is called each time the state of
     the socket is changed. This is sufficient for our purpose since we are only interested in the state
@@
             $S := S_{last}$\;
-            $E :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$)\;
-            \If{$E \wedge \text{Event\_Mask} \neq 0$}{
+            $E :=$ \lstinline[language=Ada]{Tcp_Update_Events}($S_{last}$)\;
+            \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}{
                 \KwRet{$S$}\;
@@
                 $S := S \cup S_{last}$\;
-                $E :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$)\;
-                \If{$E \wedge \text{Event\_Mask} \neq 0$}{
+                $E :=$ \lstinline[language=Ada]{Tcp_Update_Events}($S_{last}$)\;
+                \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}{
                     \KwRet{$S$}\;
@@
 
-    \textcolor{red}{\paragraph{Weakness} If two threads want to lock the same mutex at the same time, which one win?
+    \textcolor{red}{\paragraph{Weakness} If two tasks want to lock the same mutex at the same time, which one wins?
     This is probably something to discuss with Clément, because the function \texttt{tcpWaitForEvents} could have a
@@
 
+    \subsection{Bug found}
+
+    Thanks to this work, a bug has been found. The contract puts on the function \lstinline[language=Ada]{Tcp_Change_State}
+    warned of an unauthorized transition. We reproduce the code that leads to this transition and we will explain
+    why it is a bug.
+    \begin{lstlisting}[language=Ada, basicstyle=\ttfamily\small, numbers=left, numberstyle=\tiny, escapechar=\%]
+   case Sock.State is
+      when TCP_STATE_SYN_RECEIVED
+         | TCP_STATE_ESTABLISHED =>
+      -- Flush the send buffer
+      Tcp_Send (Sock, Buf, Ignore_Written,
+                SOCKET_FLAG_NO_DELAY, Error);
+      if Error /= NO_ERROR then
+         return;
+      end if;
+
+      -- Make sure all the data has been sent out
+      Tcp_Wait_For_Events % \label{bugProg:tcpWaitForEvents} %
+         (Sock       => Sock,
+          Event_Mask => SOCKET_EVENT_TX_DONE,
+          Timeout    => Sock.S_Timeout,
+          Event      => Event);
+
+      -- Timeout error?
+      if Event /= SOCKET_EVENT_TX_DONE then
+         Error := ERROR_TIMEOUT;
+         return;
+      end if;
+      % \label{bugProg:tcpWaitForEventsPost} %
+      -- Send a FIN segment
+      Tcp_Send_Segment
+         (Sock         => Sock,
+          Flags        => TCP_FLAG_FIN or TCP_FLAG_ACK,
+          Seq_Num      => Sock.sndNxt,
+          Ack_Num      => Sock.rcvNxt,
+          Length       => 0,
+          Add_To_Queue => True,
+          Error        => Error);
+      -- Failed to send FIN segment?
+      if Error /= NO_ERROR then
+         return;
+      end if;
+
+      -- Switch to the FIN-WAIT-1 state
+      Tcp_Change_State (Sock, TCP_STATE_FIN_WAIT_1);
+    \end{lstlisting}
+
+    The function \lstinline[language=Ada]{Tcp_Send} will change the state of the socket for
+    the state \ESTABLISHED{} or \CLOSEWAIT{} in our case, if no error happens. Moreover the next
+    function at line \ref{bugProg:tcpWaitForEvents} has a contract that contains:
+    \begin{lstlisting}[language=Ada, basicstyle=\ttfamily\small]
+(if (Event_Mask and SOCKET_EVENT_TX_DONE) /= 0 then
+   (if Event = SOCKET_EVENT_TX_DONE then
+      -- RST segment received
+      Model(Sock) = (Model(Sock)'Old with delta
+         S_State => TCP_STATE_CLOSED,
+         S_Reset_Flag => True) or else
+      (if Sock.S_State'Old = TCP_STATE_CLOSE_WAIT then
+         Model(Sock) = Model(Sock)'Old or else
+      elsif Sock.S_State'Old = TCP_STATE_ESTABLISHED then
+         Model(Sock) = Model(Sock)'Old or else
+         Model(Sock) = (Model(Sock)'Old with delta
+            S_State => TCP_STATE_CLOSE_WAIT))))
+    \end{lstlisting}
+    which means that at line \ref{bugProg:tcpWaitForEventsPost}, the socket can either be
+    in state \ESTABLISHED{}, \CLOSEWAIT{} or \CLOSED{} if the connection had been reset by the remote
+    during the call of \lstinline[language=Ada]{Tcp_Wait_For_Events}, when the mutex was released.
+    The precondition of the function \lstinline[language=Ada]{Tcp_Change_State} encodes the allowed
+    transitions that we can see in the automaton in \ref{fig:TCPAutomaton}, and thus the transition
+    \CLOSEWAIT{} $\rightarrow$ \FINWAITONE{} is not allowed, nor the transition \CLOSED{} $\rightarrow$ \FINWAITONE{}.
+
+    This incorrect behavior has been detected thanks to the use of Klee and the improvement
+    on how function's closure are computed. The contract for the call
+    \begin{lstlisting}[language=Ada, basicstyle=\ttfamily\small]
+    Tcp_Wait_For_Events (Sock, SOCKET_EVENT_TX_READY,
+                         Sock.S_Timeout, Event);
+    \end{lstlisting}
+    was not correct at first, because it was written by hands and thus not enough precise.
+    Then, the possible state at the end of the function \lstinline[language=Ada]{Tcp_Change_State}
+    was reduce to \ESTABLISHED{} and the same problem existed for the call
+    at line \ref{bugProg:tcpWaitForEvents}. At this point no problem was detected with SPARK,
+    because the contracts were true. With Klee, some transitions that were not considered at first,
+    because hide in subfunctions have been discovered. In particular there exists a transition from
+    \SYNRECEIVED{} to \CLOSEWAIT{} if a FIN flags is present in the same segment of the ACK that was not
+    spotted before. Putting these information in the function \lstinline[language=Ada]{Tcp_Process_One_Segment},
+    and proving the post-condition of the function \lstinline[language=Ada]{Tcp_Wait_For_Events}
+    directly with SPARK using algorithm \ref{algo:waitForEvents}, rather than not proving it like
+    it was done before, helps to correct the contracts and find the bug.
+
     \section{Future work}
 
-    Other improvements that could be done in the future:
+    Other improvements that could be done in the future (this is a not exhaustive aggregate of ideas,
+    this list is subject to change):
     \begin{itemize}
         \item Fully translate the file \texttt{tcp\_misc.c} in Ada to have more guarantees.
-        \item Same for the file \texttt{tcp\_fsm.c} that has been investigate with klee. But
+        \item Same for the file \texttt{tcp\_fsm.c} that has been investigated with klee. But
         it is not enough and we need more guarantees for this file, in order to prove the
-        interactions between threads first, and also to prove the correctness of this part
+        interactions between tasks first, and also to prove the correctness of this part
         of code w.r.t. the TCP norm.
+        \item The translation only considers one possible preprocessing. A future work would
+        be considering all the possible preprocessing by adding constants like in the C
+        version. An harder work is to find a way to keep a coherence between the C preprocessing
+        and the Ada preprocessing. It can be achieved with an automatic tool, for example.
+
+        \item It is not checked that the correct flags have been sent before changing of state.
+        It could be added by using ghost variables, or abstract states.
     \end{itemize}

### 2020-07-27 report/report.tex
Update report according to Yannick's comments

@@
         \draw ($(listen.south) + (-1em,0)$) 
-        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-2em, yshift=1em]{SYN/SYN + ACK} 
+        edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-2em, yshift=1em]{SYN/(SYN \& ACK)} 
         ($(synRcvd.east) + (0,1em)$);
@@
         \draw ($(synSent.west) + (0,-1em)$) 
-        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=3em, yshift=-1em]{SYN + ACK/ACK} 
+        edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=3em, yshift=-1em]{(SYN \& ACK)/ACK} 
         ($(established.north) + (1em,0)$);
@@
         \draw (finWait1.south east) 
-        edge[stateEdge] node[edgeLabel, xshift=1em, yshift=2em, text width=3em]{FIN + ACK/ACK} 
+        edge[stateEdge] node[edgeLabel, xshift=1em, yshift=2em, text width=3em]{(FIN \& ACK)/ACK} 
         (timeWait.north west);
@@
     We present a scenario on how a connection can be established between two TCPs,
-    when one TCP (TCP A) is in the state \CLOSED{} and the other (TCP B) is in the state \LISTEN.
+    when one TCP (TCP $\alpha$) is in the state \CLOSED{} and the other (TCP $\beta$) is in the state \LISTEN.
     \begin{enumerate}
-        \item TCP A wants to initiate a connection with TCP B. It sends a SYN segment to
-            TCP B and passes to the state \SYNSENT{}.
+        \item TCP $\alpha$ wants to initiate a connection with TCP $\beta$. It sends a SYN segment to
+            TCP $\beta$ and passes to the state \SYNSENT{}.
 
-        \item TCP B receives a segment SYN. It has to respond to this segment by a segment
+        \item TCP $\beta$ receives a segment SYN. It has to respond to this segment by a segment
             containing the flags SYN and ACK (acknowledgment of the received segment).
@@
 
-        \item When TCP A receives the segment sent by TCP B containing the flags SYN and ACK,
+        \item When TCP $\alpha$ receives the segment sent by TCP $\beta$ containing the flags SYN and ACK,
             it only has to send back an ACK of this segment to pass in the state \ESTABLISHED.
 
-        \item TCP B receives the ACK of its segment and pass in state \ESTABLISHED{} too.
+        \item TCP $\beta$ receives the ACK of its segment and pass in state \ESTABLISHED{} too.
     \end{enumerate}
     At the end of this procedure, the two TCPs are in state \ESTABLISHED{} and can start to
-    send data. Figure \ref{fig:handshake} illustrates the procedure described above. This procedure to open a connection is known as the ``three-way
-    handshake'' in the TCP norm.
+    send data. Figure \ref{fig:handshake} illustrates the procedure described above.
+    This procedure to open a connection is known as the ``three-way handshake'' in the TCP norm.
 
@@
         \begin{tikzpicture}
-            \node (A) at (0, 5) {\bfseries TCP A};
-            \node (B) at (5, 5) {\bfseries TCP B};
+            \node (A) at (0, 5) {\bfseries TCP $\alpha$};
+            \node (B) at (5, 5) {\bfseries TCP $\beta$};
             \draw[dashed] (-.7,4.8) -- (5.7,4.8);
@@
         \par
-        \caption{Three-way hanshake procedure. The solid lines represent the transition between
+        \caption{Three-way hanshake procedure. The solid lines represent the transitions between
         the states and the dashed lines represent the messages sent.}
@@
 
-    Finally, all the results found by KLEE have been reported as a post-condition of the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
+    Finally, all the results found by KLEE have been reported as a post-condition of the function
+    \lstinline[language=Ada]{Tcp_Process_One_Segment}.
     This function is essential for all the verification of concurrent parts and is reused everywhere
@@
 
-    This incorrect behavior has been detected thanks to the use of Klee and the improvement
+    This incorrect behavior has been detected thanks to the use of KLEE and the improvement
     on how function's closure are computed. The contract for the call
@@
     at line \ref{bugProg:tcpWaitForEvents}. At this point no problem was detected with SPARK,
-    because the contracts were true. With Klee, some transitions that were not considered at first,
+    because the contracts were true. With KLEE, some transitions that were not considered at first,
     because hide in subfunctions have been discovered. In particular there exists a transition from
@@
         \item Fully translate the file \texttt{tcp\_misc.c} in Ada to have more guarantees.
-        \item Same for the file \texttt{tcp\_fsm.c} that has been investigated with klee. But
+        \item Same for the file \texttt{tcp\_fsm.c} that has been investigated with KLEE. But
         it is not enough and we need more guarantees for this file, in order to prove the

### 2020-09-01 paper/paper.tex
first parts of the paper

\documentclass[runningheads]{llncs}

\usepackage{listings}
\usepackage[colorlinks]{hyperref}

\usepackage{xcolor}

\lstset{
    basicstyle=\fontsize{9}{9}\selectfont\ttfamily,
    keywordstyle=\color{blue!50}\bfseries
}

\usepackage{subcaption}

\title{Verification of a TCP/IP stack}

\author{Guillaume Cluzel\inst{1,2} \and
        Yannick Moy\inst{1} \and
        Clément Zeller\inst{3}}

        \institute{Adacore \and
        École Normale Supérieure de Lyon \and
        Oryx Embedded}

\begin{document}

\authorrunning{G. Cluzel et al.} %, Y. Moy and C. Zeller}

\maketitle

\begin{abstract}
    The abstract should briefly summarize the contents of the paper in
    150--250 words.

    \keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}

\section{Introduction}

    In the development of the Internet, there is a protocol that has a very special place: TCP.
    It has been described by Vincent Cerf and Bob Kahn who can be considered as the fathers for the Internet
    for their work, and the protocol has later been standardized in a RFC norm \cite{rfc793}.
    Since this date, the TCP has been widely used on the Internet as the underlying protocol for application layers such as HTTP,
    FTP or TLS. TCP uses itself an underlying protocol to work, the Internet layer IP protocol which is itself
    based on a link layer, such as Ethernet. This stacking of protocols is referred to as the TCP/IP stack.

    The last decade has seen a growing interest for security and privacy in all the fields where computers are
    involved. The development of safe applications thanks to static analyzers and formal methods, that was first
    restricted to critical embedded software with the Astrée analyzer~\cite{cousot2005astree} or the B method, are
    now adopted on a larger scale. New tools have been developed to promote the use of formal methods. Among them we
    can mention the SMT-solvers Z3~\cite{de2008z3} and Alter-Go~\cite{conchon:hal-01960203},
    the higher level language F*~\cite{mumon} that can prove assertions
    using these SMT-solvers or the more automatic static analyzer by abstract interpretation Frama-C~\cite{cuoq2012frama}
    that can help to identify undefined behaviors in C programs. Theorem provers, such as Coq or Isabelle, are another class
    of tools, where proofs are done manually.

    In the domain of networking and cryptography, these tools have been used to make secure implementations. 
    Recently, Reiher and al have proposed a domain-specific language, RecordFlux \cite{Reiher2019RecordFluxFM},
    to specify the message formats for real world protocols, that is written in SPARK.
    This is also the goal followed by the Everest project~\cite{bhargavan2017everest}.
    The authors implemented cryptographic functions such as AES or
    SHA2 as well as TLS, the cryptographic protocol that provides security
    to HTTPS and is its heart. The implementation is written
    in F* and is a combination of several sub-project such as miTLS~\cite{bhargavan2013implementing}
    or HACL*~\cite{zinzindohoue2017hacl}. The weakness of the Everest project is that it needs to use the untrusted
    transport layer TCP. If a bug exists in the transport layer, it can jeopardize the higher level layer.

    The importance of the TCP protocol in the network communications has driven some people to work
    on it and try to ensure the safety of the protocol. Among these works, in~\cite{smith1996formal}
    the author has tried to formally verify that a TCP implementation verify some specifications.
    Another work comes from Tom Ridge and al \cite{ridge2008rigorous} who have implemented the TCP
    protocol in the HOL proof assistant.
    But only few works has been done. Doing formal verification on a TCP implementation face up to two
    problems. The first one is the complexity of the TCP protocol that is written to be concurrent.
    The second problem comes from the norm itself that is written in a natural, informal language: English.
    Then a lot of part are under specified and extracted strong specifications can be hard.
    As a result, it does not exists a safe implementation for TCP that can be used in the industry.

    Our approach has been to consider an existing implementation of a TCP/IP stack and replace and integrate
    some bricks by Ada/SPARK code.
    We have chosen the Oryx Embedded's library, CycloneTCP\footnote{\url{https://www.oryx-embedded.com/products/CycloneTCP.html}}.
    The library is written in C, which is suitable to interface with Ada/SPARK, and is developed to run
    on embedded processors. We focused our work on rewriting in SPARK the TCP protocol because of the importance
    it has in the network communication as well as the socket API to provide a safe interface to the programmers
    to use our safe TCP implementation.

    In the first section we explore more in details the TCP protocol and the state machine that governs the behavior of the connection.
    Then in the next section we explain how we have formalized and implemented in SPARK the TCP user functions and the socket API.
    Finally in the last section we present some results of our work, like the bugs that have been found and eliminated thanks to our
    new implementation.

\section{TCP protocol and CycloneTCP library}
\label{sec:TCP}

    In this part we should present:
    \begin{itemize}
        \item The TCP protocol
        \begin{itemize}
            \item Different states
            \item Flags that can be received (mainly SYN, ACK, FIN \& RST)
            \item State machine
            \item User functions
            \item Precise that most of transitions are done when a message is received and then, outside of the user functions.
        \end{itemize}
        \item The CycloneTCP library. Present the different caracteristics, in particular
        \begin{itemize}
            \item The task model adopted to be run on a RT OS. (A task dedicated to user task, another to timers, and the last to incoming messages processing).
            \item The socket API that is provided to use UDP and TCP protocols.
        \end{itemize}
    \end{itemize}

    Any other suggestion? (2/3 pages max).

\section{Verification of the TCP protocol and the Socket API}

    Our work is focused on the key elements of the CycloneTCP library: the sockets and TCP protocol itself.
    This two parts have been replaced by SPARK code and the absence of run-time errors has been proved for
    these parts.
    However, to be integrated with the existing library, it has been necessary to reuse existing C code
    as an untrested black box, for which we cannot have strong guarantees. Thus, others techniques
    and tools have helped the verification process. They are presented in this section.

\subsection{The SPARK programming language}

    SPARK \cite{mccormick2015building} is a tool that can be used to formally prove properties on a subset of Ada such that
    information flow, absence of runtime errors and functional correctness. The subset of Ada that can be analyzed by SPARK
    grows constantly and rapidely, recently with the addition of pointers \cite{dross2020recursive}.
    In Ada, contracts can be attached to subprograms as pre- and postconditions which allows functional and modular analysis.
    The pre- and postconditions are exported to be proved by SMT solvers thanks to the Why3 tool.

\subsection{Hardening the user's socket API}

\begin{figure}[t]
\begin{subfigure}{\textwidth}
\begin{lstlisting}[language=C, frame=bottomline]
Socket *socketOpen(uint_t type, uint_t protocol);
error_t socketConnect(Socket *socket, IpAddr *remoteIpAddr,
                      uint16_t remotePort);
\end{lstlisting}
\caption{C interface.}
\label{Fig:socketInterface:C}
\end{subfigure}
\begin{subfigure}{\textwidth}
\begin{lstlisting}[language=Ada, frame=bottomline]
type Socket_Struct is private;
type Socket is access Socket_Struct;
subtype Not_Null_Socket is not null Socket;
type Socket_Type is (SOCKET_TYPE_STREAM, SOCKET_TYPE_DGRAM);
type Socket_Proto is (SOCKET_IP_PROTO_TCP, SOCKET_IP_PROTO_UDP);

procedure Socket_Open
    (Sock       :    out Socket;
     S_Type     : in     Socket_Type;
     S_Protocol : in     Socket_Protocol)
with 
    Post => (if Sock /= null then
        Sock.S_Type = S_Type and then
        not Is_Initialized_Ip(Sock.S_Remote_Ip_Addr) and then
        not Is_Initialized_Ip(Sock.S_localIpAddr));

procedure Socket_Connect
    (Sock           : in out Not_Null_Socket;
     Remote_Ip_Addr : in     IpAddr;
     Remote_Port    : in     Port;
     Error          :    out Error_T)
with
    Pre => Is_Initialized_Ip (Remote_Ip_Addr) and then
           Remote_Port > 0,
    Post => (if Error = NO_ERROR then
            Is_Initialized_Ip(Sock.S_localIpAddr) and then
            Sock.S_localIpAddr = Remote_Ip_Addr);
\end{lstlisting}
\caption{SPARK interface.}
\label{Fig:socketInterface:SPARK}
\end{subfigure}
\caption{Comparison between the C and the SPARK interface for \texttt{Socket\_Open} et \texttt{Socket\_Connect}.}
\label{Fig:socketInterface}
\end{figure}

    The socket API is a set of functions used to communicate with the protocol stack. These are therefore the functions used
    by programmers in their programs to use network protocols. However, one of the most significant problems pointed to by the
    primary author of the CycloneTCP library is the incorrect usage of the library's API. More specifically, the library's users
    tend to call the API functions with wrong arguments and they tend to forget to forget to check the value of the return code of
    function call before continuing the processing. This can lead to incorrect program behaviors when an error happen in a function
    and it can be difficult to debug.

    In order to design a better socket API, we have taken advantage of the strong-typing characteristic of Ada,
    and the features of the SPARK language, in particular the functional specification.

\subsubsection{Using types to carry information}

    Strong-typing is a first mechanism to ensure correctness of program and exclude a range of programing errors. Meaningful names for
    variables and types give a hint about the subprogram behavior, but also the valid input values accepted.
    Figure \ref{Fig:socketInterface} is a comparison between the interface of the C functions of the original library
    \lstinline{socketOpen} and \lstinline{socketConnect} and their SPARK version \lstinline{Socket_Open} and
    \lstinline{Socket_Connect}. The C function \lstinline{socketOpen} takes two arguments, the type of the socket and the protocol,
    both as unsigned integer and it returns a new socket in case of success.
    Its SPARK equivalent takes its two arguments, type and protocol, as enumeration values, which ensures correct calls
    of this function. Indeed, all the unsigned values that do not correspond to any protocol or any type of socket are eliminated.

    In SPARK, types can be constrained by a valid range of values. Another kind of constraints on types can appear in
    access types. The example on figure \ref{Fig:socketInterface:SPARK} shows a such mechanism for the type \lstinline{Not_Null_Socket},
    that is a socket with a constraint that states that the pointer has to be not null. The function \lstinline{socketConnect},
    used to connect to a remote server, requires a socket not null. While the C function contains defensive code to ensure that
    the socket passed as first argument is not null, \lstinline{Socket_Connect} directly uses information on the type of the argument
    \lstinline{Sock} to know that the socket is not null. Moreover, this information can be checked statically with GNATprove.
    Then removing the defensive code has two interests: first, it reduces the size of the program by removing lines of code,
    and secondly it reduces the number of instructions to execute, and then the global execution time.

\subsubsection{Order of the functions call}

    The TCP connection follows an order of execution as seen in the section \ref{sec:TCP}: the connection is first opened,
    then the data are sent and received, and finally the connection can be closed. Since the sockets are an interface to use
    network protocols, they follow the same order of execution. Thus, an order can be enforced thanks to the pre- and postconditions.
    If a function $f_1$ must be executed before a function $f_2$, the strategy is to add a postconditions on the socket on $f_1$
    that is also a precondition for $f_2$. Then, for the precondition of $f_2$ to be true, the function $f_1$ must have be called first.
    Such postcondition is presented on figure \ref{Fig:socketInterface:SPARK} for the function \lstinline{Socket_Connect}. After its
    execution, the field \lstinline{S_Remote_Ip_Addr} of the socket \lstinline{Sock}. Then, since all the functions to send or receive
    data and to close the connection require a connection established, these functions have a precondition that contains a condition
    that states that the condition is established, which can be written as
    \begin{lstlisting}
Pre => Is_Ip_Initialized(Sock.S_Remote_Ip_Addr);
    \end{lstlisting}
    and because the function \lstinline{Socket_Connect} is the only one that can initialize this value, the order can be respected
    only with valid code.

    The same mechanism is used with the functions \lstinline{Socket_Open} that create a valid socket when the function
    succeeds and \lstinline{Socket_Connect} that require a valid not null socket.

    GNATprove can prove that the order of the calls is respected, and then it can invalidate an incorrect program without requiring any
    tests. It helps to programmers to write correct programs.

\subsubsection{Checking error codes}

    Most of the functions of the socket API can fail. To know if the function has failed or not, and error code is returned.
    It must be checked by the caller before continuing processing at the risk of leading to an incorrect state.
    In the previous paragraph we explained that the functions that send or receive data check in their precondition if the connection
    has been initialized or not, by checking the field \lstinline{S_Remote_Ip_Addr}, and we explained that this value is initialized
    by \lstinline{Socket_Connect}. The presence in the postcondition of \lstinline{Socket_Connect} or if statement
    \lstinline{(if Error = NO_ERROR then ...)} makes that \lstinline{S_Remote_Ip_Addr} is initialized only when no error happen during
    the execution of the function. Thanks to the mechanism exposed in the previous section the next function can only be called
    if it is known that no error happen previously.

    Let \lstinline{Socket_Send} be the function to send data. The following example of code is incorrect because if \lstinline{Error} is
    different to \lstinline{NO_ERROR}, nothing ensures that \lstinline{S_Remote_Ip_Addr} is initialized.
    \begin{lstlisting}
Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
-- GNATprove: medium: precondition might fail.
Socket_Send(Sock, Data, Written, Flags, Error);
    \end{lstlisting}
    In comparison this second example of code produces no error because when \lstinline{Socket_Send} is called, GNATprove knows that
    \lstinline{ERROR} is equal to \lstinline{NO_ERROR} and then \lstinline{S_Remote_Ip_Addr} is initialized.
    \begin{lstlisting}
Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
if Error = NO_ERROR then
    Socket_Send(Sock, Data, Written, Flags, Error);
end if;
    \end{lstlisting}

    The same mechanism can be observed on figure \ref{Fig:socketInterface:SPARK}. The user has to test if the socket returned by
    \lstinline{Socket_Open} is null or not before calling \lstinline{Socket_Connect}, otherwise GNATprove will emit a message.

    Thanks to this use of GNATprove, one can ensure the correct use of the socket API, by ensuring a correct order for the
    calls of the functions and ensuring that error codes are checked.

\subsection{Conformance to the TCP protocol}

    The user functions of the TCP protocol has been rewritten in SPARK. Besides proving the absence of run-time error in these functions,
    it has been possible to check that some properties of the TCP protocol are not violated in the code.

\subsubsection{Extracting a specification from the TCP norm}

    Extracting a specification is mandatory to provide later functional contracts to the user functions.
    That can be achieved by reading the norm. However, the norm is underspecified in many cases and only one possible implementation
    is depicted. Thus, the specifications we have extracted are conformant with the norm but they differ a little to
    what can be found in the RFC.

    Among the specification we have extracted:
    \begin{itemize}
        \item The transitions between the states must respect the order given by the state machine described
              in figure \ref{Fig:statemachie}.
              In particular, a transition from a state to another is valid only if there exists a transition in the state machine
              between these two states.
        \item User functions contains functional specifications in the norm, that describe what have to be done when the function
              is called depending on the state of the socket~\cite[p.~52]{rfc793}.
              In particular, a socket in some state will return an error for certain calls of user functions. Figure \ref{Fig:constraint}
              lists all the constaint on the state to call a function, and the resulting states.
    \end{itemize}

\subsubsection{Rewritting TCP user functions}

    The user functions have been translated in SPARK to prove the absence of run-time errors in them. The low level IP functions
    that are called are assumed to be correct and then are used as an untrested black box. They are supposed to not interact with TCP
    special fields. However, there exists another kind of subprograms that interact directly and can modify the state of the socket
    during the execution of the program: the procedure \lstinline{Tcp_Wait_For_Events}. This procedure is called when the code needs
    to wait for an actions to be executed by another thread. This action can be, for example, the reception of a segment. But when a
    segment is received, the socket can change of state. Then it is necessary to know what are the exact possible states when the
    procedure \lstinline{Tcp_Wait_For_Events} returns if we wants to know that there is no violation of the norm in the further transitions.

    \begin{itemize}
        \item En revanche certains autres appels dans le code \lstinline{Tcp_Wait_For_Events} attendent qu'un autre thread qui exécutent des actions sur TCP ait changé l'état de la socket pour retourner.
        \item Ces évènements sont donc à prendre en compte. Il faut savoir précisément quel est l'état de la socket quand la fonction retourne si on veut calculer de manière précise les états à la sortie de la fonction, et le fait qu'il n'y ait pas de violation de la norme lors des transitions.
        \item Il faut donc analyser ce que ces autres fonctions font pour leur ajouter des contrats corrects et ainsi prouver que les contrats qu'on avait exatraits sont bien vérifiés.
        \item Le problème est que SPARK n'a aucun moyen pour prendre en compte la concurrence de manière native. Il faut donc utiliser un moyen détourné, manuel, pour simuler ce qui peut être exécuté par les autres tasks.
    \end{itemize}

\subsubsection{The use of KLEE to extract contracts}

    Symbolic execution is a means to execute a program using symbolic variables instead of concrete values. It accumulates contraints
    on those values which are used to generate proof obligations. Symbolic execution has been successfully used with other verification
    methods. Indeed, in~\cite{vanoverberghe2008using} the authors present how to use symbolic execution to improve deductive verification,
    and in~\cite{kassios2012comparing} the authors explain that symbolic execution can be good at infering contracts.
    For our work, KLEE has been used. It is built on top of LLVM as it takes LLVM bytecode in input
    and works well with C code thanks to a very simple interface.

\section{Results and bugs found}

    \begin{itemize}
        \item Bug de concurrence -> SPARK
        \item Memory leak -> SPARK
        \item Problèmes trouvés.
        \item Résultats sur la nouvelle implémentation (performance, taille du code).
        \item Activation des préconditions pour voir si elles sont toutes valides.
    \end{itemize}

\section{Conclusion}

\bibliographystyle{splncs04}
\bibliography{biblio}

\end{document}

### 2020-09-01 paper/paper.tex
corrections

@@
     It has been described by Vincent Cerf and Bob Kahn who can be considered as the fathers for the Internet
-    for their work, and the protocol has later been standardized in a RFC norm \cite{rfc793}.
+    for their work, and the protocol has later been standardized in a RFC norm~\cite{rfc793}.
     Since this date, the TCP has been widely used on the Internet as the underlying protocol for application layers such as HTTP,
@@
     In the domain of networking and cryptography, these tools have been used to make secure implementations. 
-    Recently, Reiher and al have proposed a domain-specific language, RecordFlux \cite{Reiher2019RecordFluxFM},
+    Recently, Reiher and al have proposed a domain-specific language, RecordFlux~\cite{Reiher2019RecordFluxFM},
     to specify the message formats for real world protocols, that is written in SPARK.
@@
     in F* and is a combination of several sub-project such as miTLS~\cite{bhargavan2013implementing}
-    or HACL*~\cite{zinzindohoue2017hacl}. The weakness of the Everest project is that it needs to use the untrusted
-    transport layer TCP. If a bug exists in the transport layer, it can jeopardize the higher level layer.
+    or HACL*~\cite{zinzindohoue2017hacl}. The weakness of the Everest project lies in its need to use the untrusted
+    transport layer TCP. If a bug exists in the transport layer, it can also jeopardize the higher level layers.
 
@@
     the author has tried to formally verify that a TCP implementation verify some specifications.
-    Another work comes from Tom Ridge and al \cite{ridge2008rigorous} who have implemented the TCP
+    Another work comes from Tom Ridge and al~\cite{ridge2008rigorous} who have implemented the TCP
     protocol in the HOL proof assistant.
@@
 
+    Figures à inclures ici:
+    \begin{itemize}
+        \item Une figure qui présente l'automate de transition. Expliquer que certaines transition sont effectuées par une action de l'utilisateur,
+              et que d'autres sont effectuées quand un segment est reçu.
+        \item Une seconde figure qui présente l'organisation d'une pile TCP/IP, avec les différentes couches et les différents protocols.
+              Ça permettra de présenter en même temps les parties qui ont été vérifiées formellement, et les parties de la pile existante qui
+              sont réutilisées. Ça permet de montrer la frontière, et les points critiques. Notamment le point critique qui a nécessité l'utilisation
+              de KLEE.
+    \end{itemize}
+
 \section{Verification of the TCP protocol and the Socket API}
@@
     Our work is focused on the key elements of the CycloneTCP library: the sockets and TCP protocol itself.
-    This two parts have been replaced by SPARK code and the absence of run-time errors has been proved for
+    These two parts have been replaced by SPARK code and the absence of run-time errors has been proved for
     these parts.
     However, to be integrated with the existing library, it has been necessary to reuse existing C code
-    as an untrested black box, for which we cannot have strong guarantees. Thus, others techniques
+    as an untrested black box, for which we cannot have strong guarantees. Thus, other techniques
     and tools have helped the verification process. They are presented in this section.
+    The source code of the reworked TCP/IP stack can be found on Github\footnote{\url{https://github.com}}.
 
@@
 
-    SPARK \cite{mccormick2015building} is a tool that can be used to formally prove properties on a subset of Ada such that
+    SPARK~\cite{mccormick2015building} is a tool that can be used to formally prove properties on a subset of Ada such that
     information flow, absence of runtime errors and functional correctness. The subset of Ada that can be analyzed by SPARK
-    grows constantly and rapidely, recently with the addition of pointers \cite{dross2020recursive}.
+    grows constantly and rapidly, recently with the addition of pointers~\cite{dross2020recursive}.
     In Ada, contracts can be attached to subprograms as pre- and postconditions which allows functional and modular analysis.
@@
     primary author of the CycloneTCP library is the incorrect usage of the library's API. More specifically, the library's users
-    tend to call the API functions with wrong arguments and they tend to forget to forget to check the value of the return code of
-    function call before continuing the processing. This can lead to incorrect program behaviors when an error happen in a function
+    tend to call the API functions with wrong arguments and they tend to forget to check the value of the return code of the
+    function call before continuing the processing. This can lead to incorrect program behaviors when an error occurs in a function
     and it can be difficult to debug.
@@
     In order to design a better socket API, we have taken advantage of the strong-typing characteristic of Ada,
-    and the features of the SPARK language, in particular the functional specification.
+    and the features of the SPARK language, in particular the functional specifications.
 
@@
 
-    Strong-typing is a first mechanism to ensure correctness of program and exclude a range of programing errors. Meaningful names for
+    Strong-typing is a first mechanism to ensure correctness of programs and exclude a range of programing errors. Meaningful names for
     variables and types give a hint about the subprogram behavior, but also the valid input values accepted.
@@
     \lstinline{Socket_Connect}. The C function \lstinline{socketOpen} takes two arguments, the type of the socket and the protocol,
-    both as unsigned integer and it returns a new socket in case of success.
+    both as unsigned integers and it returns a new socket in case of success.
     Its SPARK equivalent takes its two arguments, type and protocol, as enumeration values, which ensures correct calls
@@
 
-    In SPARK, types can be constrained by a valid range of values. Another kind of constraints on types can appear in
+    In SPARK, types can be constrained by a valid range of values. Another kind of constraint on types can appear in
     access types. The example on figure \ref{Fig:socketInterface:SPARK} shows a such mechanism for the type \lstinline{Not_Null_Socket},
@@
     network protocols, they follow the same order of execution. Thus, an order can be enforced thanks to the pre- and postconditions.
-    If a function $f_1$ must be executed before a function $f_2$, the strategy is to add a postconditions on the socket on $f_1$
-    that is also a precondition for $f_2$. Then, for the precondition of $f_2$ to be true, the function $f_1$ must have be called first.
+    If a function $f_1$ must be executed before a function $f_2$, the strategy is to add a postcondition on the socket on $f_1$
+    that is also a precondition for $f_2$. Then, for the precondition of $f_2$ to be true, the function $f_1$ must have been called first.
     Such postcondition is presented on figure \ref{Fig:socketInterface:SPARK} for the function \lstinline{Socket_Connect}. After its
@@
     \end{lstlisting}
-    and because the function \lstinline{Socket_Connect} is the only one that can initialize this value, the order can be respected
-    only with valid code.
+    and because the function \lstinline{Socket_Connect} is the only one that can initialize this value, the order is respected
+    if and only if the code is valid.
 
@@
     GNATprove can prove that the order of the calls is respected, and then it can invalidate an incorrect program without requiring any
-    tests. It helps to programmers to write correct programs.
+    tests. It helps programmers to write correct programs.
 
@@
     by \lstinline{Socket_Connect}. The presence in the postcondition of \lstinline{Socket_Connect} or if statement
-    \lstinline{(if Error = NO_ERROR then ...)} makes that \lstinline{S_Remote_Ip_Addr} is initialized only when no error happen during
+    \lstinline{(if Error = NO_ERROR then ...)} makes that \lstinline{S_Remote_Ip_Addr} is initialized only when no error occurs during
     the execution of the function. Thanks to the mechanism exposed in the previous section the next function can only be called
-    if it is known that no error happen previously.
+    if it is known that no error has occurred previously.
 
@@
         \item The transitions between the states must respect the order given by the state machine described
-              in figure \ref{Fig:statemachie}.
+              in figure \ref{Fig:statemachine}.
               In particular, a transition from a state to another is valid only if there exists a transition in the state machine
@@
     during the execution of the program: the procedure \lstinline{Tcp_Wait_For_Events}. This procedure is called when the code needs
-    to wait for an actions to be executed by another thread. This action can be, for example, the reception of a segment. But when a
+    to wait for an events to become true. The event waited can become true, for example, when a segment is received. But when a
     segment is received, the socket can change of state. Then it is necessary to know what are the exact possible states when the
-    procedure \lstinline{Tcp_Wait_For_Events} returns if we wants to know that there is no violation of the norm in the further transitions.
+    procedure \lstinline{Tcp_Wait_For_Events} returns if we wants to be sure that there is no violation of the norm in the further transitions.
+
+    SPARK does not have a very advanced native way to deal with concurrency. It is not possible for the tool to infer precisely the
+    interferences of other tasks. Thus, using SPARK is not the right way to do, and a manual solution has been considered. It consists
+    of analysing what is done in other tasks. Then a contract can be add to \lstinline{Tcp_Wait_For_Events} that describes the possible
+    interractions of the other tasks during the wait.
 
@@
         \item En revanche certains autres appels dans le code \lstinline{Tcp_Wait_For_Events} attendent qu'un autre thread qui exécutent des actions sur TCP ait changé l'état de la socket pour retourner.
-        \item Ces évènements sont donc à prendre en compte.
-Il faut savoir précisément quel est l'état de la socket quand la fonction retourne si on veut calculer de manière précise les états à la sortie de la fonction, et le fait qu'il n'y ait pas de violation de la norme lors des transitions.
         \item Il faut donc analyser ce que ces autres fonctions font pour leur ajouter des contrats corrects et ainsi prouver que les contrats qu'on avait exatraits sont bien vérifiés.

### 2020-09-02 paper/paper.tex
Continue the explainations

@@
 \usepackage{xcolor}
+
+\usepackage[ruled]{algorithm2e}
+\usepackage{amsmath,amsfonts}
 
@@
     different to \lstinline{NO_ERROR}, nothing ensures that \lstinline{S_Remote_Ip_Addr} is initialized.
-    \begin{lstlisting}
+    \begin{lstlisting}[language=Ada]
 Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
@@
     \lstinline{ERROR} is equal to \lstinline{NO_ERROR} and then \lstinline{S_Remote_Ip_Addr} is initialized.
-    \begin{lstlisting}
+    \begin{lstlisting}[language=Ada]
 Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
@@
 
-    The user functions of the TCP protocol has been rewritten in SPARK. Besides proving the absence of run-time error in these functions,
+    The user functions of the TCP protocol have been rewritten in SPARK. Besides proving the absence of run-time error in these functions,
     it has been possible to check that some properties of the TCP protocol are not violated in the code.
@@
 
-    Among the specification we have extracted:
+    Among the specifications we have extracted:
     \begin{itemize}
@@
               In particular, a socket in some state will return an error for certain calls of user functions. Figure \ref{Fig:constraint}
-              lists all the constaint on the state to call a function, and the resulting states.
+              lists all the constraints on the state of the socket before and after the call.
     \end{itemize}
 
-\subsubsection{Rewritting TCP user functions}
+\subsubsection{Rewriting TCP user functions}
 
     The user functions have been translated in SPARK to prove the absence of run-time errors in them. The low level IP functions
-    that are called are assumed to be correct and then are used as an untrested black box. They are supposed to not interact with TCP
-    special fields. However, there exists another kind of subprograms that interact directly and can modify the state of the socket
+    that are called are assumed to be correct and then are used as an untrusted black box. They are supposed to not interact with TCP
+    special fields. However, there exists another kind of subprogram that interacts and can modify the state of the socket
     during the execution of the program: the procedure \lstinline{Tcp_Wait_For_Events}. This procedure is called when the code needs
-    to wait for an events to become true. The event waited can become true, for example, when a segment is received. But when a
-    segment is received, the socket can change of state. Then it is necessary to know what are the exact possible states when the
-    procedure \lstinline{Tcp_Wait_For_Events} returns if we wants to be sure that there is no violation of the norm in the further transitions.
+    to wait for an event to become true. The event waited can become true, for example, when a segment is received. But when a
+    segment is received, the state of state can change. Then it is necessary to know what are the exact possible states when the
+    procedure \lstinline{Tcp_Wait_For_Events} returns if we want to be sure that there is no violation of the norm in the further transitions.
 
-    SPARK does not have a very advanced native way to deal with concurrency. It is not possible for the tool to infer precisely the
-    interferences of other tasks. Thus, using SPARK is not the right way to do, and a manual solution has been considered. It consists
-    of analysing what is done in other tasks. Then a contract can be add to \lstinline{Tcp_Wait_For_Events} that describes the possible
+    SPARK does not have a very advanced native way to deal with concurrency. In particular, it is not possible for the tool to infer precisely the
+    interferences that comes from other tasks. Thus, using SPARK is not the right way to do, and a manual solution has been considered. It consists
+    of analysing what is done in other tasks. Then a contract can be added to \lstinline{Tcp_Wait_For_Events} to describe the possible
     interractions of the other tasks during the wait.
 
-    \begin{itemize}
-        \item En revanche certains autres appels dans le code \lstinline{Tcp_Wait_For_Events} attendent qu'un autre thread qui exécutent des actions sur TCP ait changé l'état de la socket pour retourner.
-        \item Il faut donc analyser ce que ces autres fonctions font pour leur ajouter des contrats corrects et ainsi prouver que les contrats qu'on avait exatraits sont bien vérifiés.
-        \item Le problème est que SPARK n'a aucun moyen pour prendre en compte la concurrence de manière native.
-Il faut donc utiliser un moyen détourné, manuel, pour simuler ce qui peut être exécuté par les autres tasks.
-    \end{itemize}
+\subsubsection{The use of KLEE to extract contracts}
 
-\subsubsection{The use of KLEE to extract contracts}
+    When a TCP segment is received, that segment is processed through a C function named \lstinline{tcpProcessSegment}.
+    Depending on the flags contains in the segment, and the state of the socket at the moment of the reception of the segment, different
+    transitions can be done. Reading the code might be enough to infer contracts connecting the state before the reception of the segment
+    and after. But from experience it is not enough and a more formal approach is necessary. Translating this function in SPARK to analyze
+    its behavior would have been a correct way to proceed, but because of a lack of time, another method has been chosen: symbolic execution.
 
@@
     and works well with C code thanks to a very simple interface.
+
+    In order to extract contracts from the C function \lstinline{tcpProcessSegment}, we have started by building a small main function that
+    simulates the reception of a segment. This function is then executed by KLEE. More precisely, first, with the KLEE intrinsic functions, we
+    create a symbolic segment (representing any segment that can be received). Secondly, the socket is changed for a particular state $S_1$. Thirdly, the
+    function is executed by KLEE. After its call, an assertion about the states $S'_1, \dots, S'_k$ of the socket after the execution
+    of the function is added to check what are the resulting states. The states $S'_1, \dots, S'_k$ are extracted from the norm and adapted to the CycloneTCP
+    implementation that can differ on some points.
+    After this work, we then have a postcondition for \lstinline{tcpProcessSegment}, that can be reported in the SPARK code for further proofs, of the form:
+    \begin{lstlisting}[language=Ada,mathescape=true]
+    if Sock.State'Old = $S_1$ then Sock.State in $S'_1$ | $S'_2$ | ... | $S'_k$
+    \end{lstlisting}
+
+\subsubsection{Solution to the concurrency challenge}
+
+    The previous part shows how we can compute all the transitions that can be done when a single message is received.
+    It is now necessary to understand how the waiting procedure \lstinline{Tcp_Wait_For_Events} works.
+    When it is called, it checks if the event is already true, thanks to a call to the function \lstinline{Tcp_Update_Events}.
+    If it is not the case, the mutex that enforces the mutual exclusion of access to the socket is unlocked.
+    Messages can be received until the event waited becomes true.
+    Once a reception makes true the event, the function \lstinline{Tcp_Wait_For_Events} locks again the mutex and returns.
+    A possible algorithm to simulate the reception of segment and compute the least set of states that can result of the wait can be done
+    by the algorithm \ref{algo:waitForEvents}.
+
+    \begin{algorithm}
+        \SetKwFunction{ProcessSegm}{\lstinline[language=Ada]{Tcp\_Wait\_For\_Events}}
+        \SetKwProg{Fn}{function}{}{end}
+        \Fn{\ProcessSegm{Socket, Event, Event\_Mask}}{
+            $S_{last} := \text{\textit{Socket}}$\;
+            $S := S_{last}$\;
+            $E :=$ \lstinline[language=Ada]{Tcp_Update_Events}($S_{last}$)\;
+            \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}{
+                \KwRet{$S$}\;
+            }
+            \For{$i=1$ \KwTo $3$}{
+                $S_{last} :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$) \;
+                $S := S \cup S_{last}$\;
+                $E :=$ \lstinline[language=Ada]{Tcp_Update_Events}($S_{last}$)\;
+                \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}{
+                    \KwRet{$S$}\;
+                }
+            }
+            \KwRet{$\emptyset$}\;
+        }
+        \caption{Function to compute the possible states after a wait for a particular event.}
+        \label{algo:waitForEvents}
+    \end{algorithm}
 
@@
 
-\section{Conclusion}
+\section{Conclusion and future work}
 

### 2020-09-03 paper/paper.tex
Typo

@@
     now adopted on a larger scale. New tools have been developed to promote the use of formal methods. Among them we
-    can mention the SMT-solvers Z3~\cite{de2008z3} and Alter-Go~\cite{conchon:hal-01960203},
+    can mention the SMT-solvers Z3~\cite{de2008z3} and Alt-Ergo~\cite{conchon:hal-01960203},
     the higher level language F*~\cite{mumon} that can prove assertions

### 2020-09-04 paper/paper.tex
More on the paper

@@
 }
+
+\usepackage{tikz}
+\usetikzlibrary{decorations.pathreplacing, patterns}
 
@@
 
+\subsection{TCP protocol}
+
+\subsection{CycloneTCP library organisation}
+
+    \input{OryxStack.tex}
+
+    The CycloneTCP library is TCP/IP stack that implements network protocols. It is particularly well adapted to embedded
+    platforms and microcontrolers. As a result, the code is written to compile on those platforms and to be used with different
+    real time OS, like for intance FreeRTOS.
+
+    Figure \ref{Fig:TcpStack} gives an overview of the different protocols supported.
+
     In this part we should present:
@@
     However, to be integrated with the existing library, it has been necessary to reuse existing C code
-    as an untrested black box, for which we cannot have strong guarantees. Thus, other techniques
-    and tools have helped the verification process. They are presented in this section.
-    The source code of the reworked TCP/IP stack can be found on Github\footnote{\url{https://github.com}}.
+    as an untrested black box, for which we cannot have strong guarantees. Figure \ref{Fig:TcpStack} gives an overview of
+    what have been translated and what have been reused. The box for TCP is hatched because only the user level functions
+    has been translated in SPARK. The parsing and processing of incoming segment is let to the C code.
+    Thus, other techniques and tools have helped the verification process, in particular for the TCP part. They are presented in this section.
+    The source code of the reworked TCP/IP stack can be found on Github\footnote{\url{https://github.com/AdaCore/Http_Cyclone}}.
 
@@
 
+    The use of SPARK for the verification of the TCP protocol has given good results. It has helped to find bugs
+    in the existing implementation. But more than finding bugs, the interest of this work lie in the face that it is
+    a big step toward a secure implementation. It has been proved by the tool GNATprove that our implementation is free
+    of run-time errors and that all the transitions in the TCP state machine are done with respect to the norm.
+
+    However weeknesses exist our implementation, in particular because all the underlying layers are still written in C.
+    The principal function of those layers is to format packet before they are sent, or parse incoming packets, check
+    their integrity and transmit the resulting payload to the corresponding upper level layer.
+    This processing part can be a source of errors and bugs as explain in~\cite{Reiher2019RecordFluxFM}.
+    Using the RecordFlux DSL to parse the packets of the different protocols is the next step to make the stack safer.
+    Using RecordFlux to parse TCP message would be a big step to finish the translation of the TCP protocol in SPARK.
+    The postconditions extracted by KLEE could be proved by GNATprove for a safer result.
+
+    Our verification of the TCP protocol only focuses on the validity of the transitions in the TCP state machine,
+    but some other properties are formulated in the TCP norm, and could be verified in the implementation to make
+    it even more robust.
+
 \bibliographystyle{splncs04}

### 2020-09-04 paper/OryxStack.tex
Figure of the Oryx Embedded stack

\newlength\hatchdistance
\hatchdistance=10pt
\pgfdeclarepatternformonly[\hatchdistance,2pt]{north east hatch}% name
    {\pgfqpoint{-1pt}{-1pt}}% below left
    {\pgfqpoint{\hatchdistance}{\hatchdistance}}% above right
    {\pgfpoint{\hatchdistance-1pt}{\hatchdistance-1pt}}%
    {
        \pgfsetcolor{orange!40}
        \pgfsetlinewidth{2pt}
        \pgfpathmoveto{\pgfqpoint{0pt}{0pt}}
        \pgfpathlineto{\pgfqpoint{\hatchdistance}{\hatchdistance}}
        \pgfusepath{stroke}
    }
\begin{figure}
    \centering\begin{tikzpicture}[x=.12\textwidth,y=.55cm,
        application/.style={font=\sffamily\scriptsize,
                            draw,
                            minimum width=.117\textwidth,
                            minimum height=.5cm},
        session/.style={font=\sffamily\scriptsize,
                        draw,
                        minimum width=.717\textwidth,
                        minimum height=.5cm},
        transport/.style={font=\sffamily\scriptsize,
                          draw,
                          minimum width=.297\textwidth,
                          minimum height=.5cm},
        network/.style={font=\sffamily\scriptsize,
                        draw,
                        minimum width=.357\textwidth,
                        minimum height=.5cm},
        datalink/.style={font=\sffamily\scriptsize,
                         draw,
                         minimum width=.141\textwidth,
                         minimum height=.5cm}
    ]
        \node[application] at (0,10) {HTTP};
        \node[application] at (1,10) {HTTP/2};
        \node[application] at (2,10) {MQTT};
        \node[application] at (3,10) {MQTT-SN};
        \node[application] at (4,10) {CoAP};
        \node[application] (CoinA) at (5,10) {FTP};
        \node[application] at (0,9)  {SMTP};
        \node[application] at (1,9)  {SNTP};
        \node[application] at (2,9)  {DNS};
        \node[application] at (3,9)  {NetBIOS};
        \node[application] at (4,9)  {SNMPv3};
        \node[application] at (5,9)  {TFTP};
        \node[application, minimum width=.237\textwidth] at (.5,8) {WebSocket};
        \node[application] at (2,8)  {mDNS};
        \node[application] at (3,8)  {DNS-SD};
        \node[application] at (4,8)  {DHCP};
        \node[application] (CoinB) at (5,8)  {DHCPv6};
        \node[session, fill=green!70!yellow!20] (Socket) at (2.5, 7) {Socket};
        \node[transport, preaction={fill, green!70!yellow!20}, pattern=north east hatch, pattern color=orange!40] at (.75,6) {TCP};
        \node[transport, fill=orange!40] at (3.25,6) {UDP};
        \node[application, fill=orange!40] (RAW) at (5,6) {RAW};
        \node[network, fill=orange!40] at (1, 5) {IPv4};
        \node[network, fill=orange!40] (IPv6) at (4, 5) {IPv6};
        \node[application, fill=orange!40] at (0,4) {ARP};
        \node[application, fill=orange!40] at (1,4) {Auto-IP};
        \node[application, fill=orange!40] at (2,4) {NDP};
        \node[application, fill=orange!40] at (3,4) {SLAAC};
        \node[application, fill=orange!40] (ICMP) at (4,4) {ICMP};
        \node[application, fill=orange!40] (MLDv1) at (5,4) {MLDv1};
        \node[datalink, fill=orange!40] at (0.10, 3) {Ethernet};
        \node[datalink, fill=orange!40] at (1.3, 3) {Wi-Fi};
        \node[datalink, fill=orange!40] at (2.5, 3) {PPP};
        \node[datalink, fill=orange!40] at (3.7, 3) {USB/RNDIS};
        \node[datalink, fill=orange!40] (G3) at (4.9, 3) {G3-PLC};

        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]CoinA.north east) -- ([xshift=2pt]CoinB.south east)
            node [midway, anchor=west, font=\sffamily\scriptsize] {Application layer};
        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]Socket.north east) -- ([xshift=2pt]Socket.south east)
            node [midway, anchor=west, font=\sffamily\scriptsize] {Session layer};
        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]RAW.north east) -- ([xshift=2pt]RAW.south east)
            node [midway, anchor=west, font=\sffamily\scriptsize] {Transport layer};
        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]IPv6.north east) -- ([xshift=2pt]MLDv1.south east)
            node [midway, anchor=west, font=\sffamily\scriptsize] {Network layer};
        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]G3.north east) -- ([xshift=2pt]G3.south east)
            node [midway, anchor=west, font=\sffamily\scriptsize] {Data Link layer};
    \end{tikzpicture}

    \caption{Overview of the TCP/IP CycloneTCP stack in the OSI model. Green parts represent
             what have been translated in SPARK and in orange the underlying untrusted part written in C.}
    \label{Fig:TcpStack}
\end{figure}

### 2020-09-04 paper/paper.tex
Rework section on SPARK inserted in new top-level section for Specification

@@
 
+\section{Specification Techniques Used}
+
+\subsection{The SPARK Programming Language}
+
+Ada is a general-purpose procedural programming language. The design of the Ada
+language puts great emphasis on the safety and correctness of the program. This
+objective is realized by using a readable syntax that uses keywords instead of
+symbols where reasonable. The type system is strong and strict and many
+potential violations of type constraints can be detected statically by the
+compiler. If not, a run-time check is inserted into the program, to guarantee
+the detection of incorrect situations. Ada 2012 introduced contract based
+programming to Ada. In particular, it is possible to attach pre- and
+postconditions to subprograms\footnote{In Ada, a distinction is made between
+  functions that return a value, and procedures, which do
+  not. \emph{Subprogram} is the term that designates both.}.  These conditions
+can be checked during the execution of the program, just like assertions.
+
+SPARK~\cite{mccormick2015building} is the name of a platform that provides
+formal verification for Ada. It uses the user-provided contracts and attempts
+to prove that the runtime checks cannot fail and that postconditions are
+established by the corresponding subprograms.  As formal verification for the
+whole Ada language would be intractable, SPARK is also the name of the subset
+of the Ada language that is supported by the SPARK tool called
+GNATprove\footnote{\url{http://docs.adacore.com/spark2014-docs/html/ug/}}.
+This subset contains almost all features of Ada, though sometimes in a
+restricted form.  In particular, expressions should be free from side effects,
+and aliasing is forbidden (no two variables should share the same memory
+location or overlap in memory), including when using pointers thanks to the use
+of an ownership policy~\cite{dross2020recursive}.  This restriction greatly
+simplifies the memory model used in the SPARK tool: any program variables can
+be reasoned about independently from other variables.
+
+GNATprove performs two kinds of analysis modularly on individual subprograms:
+information flow to detect reads of uninitialized data and violations of flow
+contracts; deductive verification based on the Why3 platform to generate
+verification conditions for SMT solvers via a weakest-precondition calculus to
+detect possible runtime errors and violations of functional contracts.
+
+\subsection{Interfacing SPARK and C Code}
+
+describe import of C subprograms/export of SPARK subprograms, and how data
+types are mapped, also contracts on imported subprograms
+
+\subsection{Dealing with Pointers}
+
+both at top-level and as fields in structures. explain how it relies on the
+ownership model. maybe a little bit on dynamic allocation/deallocation (which
+will introduce the section later on memory leaks)
+
+\subsection{Specifying the Frame Condition}
+
+describe how models of sockets are defined and used in contracts, based on
+ghost code, delta aggregates and Old attribute
+
 \section{Verification of the TCP protocol and the Socket API}
@@
     The source code of the reworked TCP/IP stack can be found on Github\footnote{\url{https://github.com/AdaCore/Http_Cyclone}}.
-
-\subsection{The SPARK programming language}
-
-    SPARK~\cite{mccormick2015building} is a tool that can be used to formally prove properties on a subset of Ada such that
-    information flow, absence of runtime errors and functional correctness. The subset of Ada that can be analyzed by SPARK
-    grows constantly and rapidly, recently with the addition of pointers~\cite{dross2020recursive}.
-    In Ada, contracts can be attached to subprograms as pre- and postconditions which allows functional and modular analysis.
-    The pre- and postconditions are exported to be proved by SMT solvers thanks to the Why3 tool.
 

### 2020-09-04 paper/paper.tex
Description of the Oryx Library

@@
 \usepackage{amsmath,amsfonts}
+
+\usepackage[utf8]{inputenc}
 
@@
         Yannick Moy\inst{1} \and
-        Clément Zeller\inst{3}}
+        Cl\'ement Zeller\inst{3}}
 
         \institute{Adacore \and
-        École Normale Supérieure de Lyon \and
+        \'Ecole Normale Sup\'erieure de Lyon \and
         Oryx Embedded}
@@
     involved. The development of safe applications thanks to static analyzers and formal methods, that was first
-    restricted to critical embedded software with the Astrée analyzer~\cite{cousot2005astree} or the B method, are
+    restricted to critical embedded software with the Astr\'ee analyzer~\cite{cousot2005astree} or the B method, are
     now adopted on a larger scale. New tools have been developed to promote the use of formal methods. Among them we
@@
 
-    The CycloneTCP library is TCP/IP stack that implements network protocols. It is particularly well adapted to embedded
-    platforms and microcontrolers. As a result, the code is written to compile on those platforms and to be used with different
-    real time OS, like for intance FreeRTOS.
+    The CycloneTCP library is TCP/IP stack that implements network protocols. 
+    Figure \ref{Fig:TcpStack} gives an overview of the different protocols supported. TCP and UDP protocols have a central place
+    in the stack. An API is provided to use those network protocols: the Socket API. It is a common way to use transport protocols,
+    and the application protocols in the stack, such as HTTP or FTP use it.
 
-    Figure \ref{Fig:TcpStack} gives an overview of the different protocols supported.
+    The stack is particularly well adapted to embedded platforms and microcontrolers. As a result, the code is written
+    to compile on those platforms and to be used with different real time OS, like for intance FreeRTOS.
+    Hence, a model based on multiple tasks has been adopted to implement the TCP protocol as defined in the norm, with,
+    in addition, complex cooperation mechanisms. A task is dedicated to the user code, another to timers, and the last is
+    assigned to the processing of incoming segments.
 
@@
               et que d'autres sont effectuées quand un segment est reçu.
-        \item Une seconde figure qui présente l'organisation d'une pile TCP/IP, avec les différentes couches et les différents protocols.
-              Ça permettra de présenter en même temps les parties qui ont été vérifiées formellement, et les parties de la pile existante qui
-              sont réutilisées. Ça permet de montrer la frontière, et les points critiques. Notamment le point critique qui a nécessité l'utilisation
-              de KLEE.
     \end{itemize}

### 2020-09-04 paper/paper.tex
Last changes

@@
 
+    %\input{Automaton.tex}
+
+In this part we should present:
+\begin{itemize}
+    \item The TCP protocol
+    \begin{itemize}
+        \item Different states
+        \item Flags that can be received (mainly SYN, ACK, FIN \& RST)
+        \item State machine
+        \item User functions
+        \item Precise that most of transitions are done when a message is received and then, outside of the user functions.
+    \end{itemize}
+\end{itemize}
+
+Figures à inclures ici:
+\begin{itemize}
+    \item Une figure qui présente l'automate de transition. Expliquer que certaines transition sont effectuées par une action de l'utilisateur,
+          et que d'autres sont effectuées quand un segment est reçu.
+\end{itemize}
+
 \subsection{CycloneTCP library organisation}
@@
     in addition, complex cooperation mechanisms. A task is dedicated to the user code, another to timers, and the last is
-    assigned to the processing of incoming segments.
-
-    In this part we should present:
-    \begin{itemize}
-        \item The TCP protocol
-        \begin{itemize}
-            \item Different states
-            \item Flags that can be received (mainly SYN, ACK, FIN \& RST)
-            \item State machine
-            \item User functions
-            \item Precise that most of transitions are done when a message is received and then, outside of the user functions.
-        \end{itemize}
-        \item The CycloneTCP library. Present the different caracteristics, in particular
-        \begin{itemize}
-            \item The task model adopted to be run on a RT OS. (A task dedicated to user task, another to timers, and the last to incoming messages processing).
-            \item The socket API that is provided to use UDP and TCP protocols.
-        \end{itemize}
-    \end{itemize}
-
-    Any other suggestion? (2/3 pages max).
-
-    Figures à inclures ici:
-    \begin{itemize}
-        \item Une figure qui présente l'automate de transition. Expliquer que certaines transition sont effectuées par une action de l'utilisateur,
-              et que d'autres sont effectuées quand un segment est reçu.
-    \end{itemize}
+    assigned to the processing of incoming segments. Then, user tasks can block without interfering the TCP/IP stack.
 

### 2020-09-07 paper/paper.tex
rework intro

@@
 
-    In the development of the Internet, there is a protocol that has a very special place: TCP.
-    It has been described by Vincent Cerf and Bob Kahn who can be considered as the fathers for the Internet
-    for their work, and the protocol has later been standardized in a RFC norm~\cite{rfc793}.
-    Since this date, the TCP has been widely used on the Internet as the underlying protocol for application layers such as HTTP,
-    FTP or TLS. TCP uses itself an underlying protocol to work, the Internet layer IP protocol which is itself
-    based on a link layer, such as Ethernet. This stacking of protocols is referred to as the TCP/IP stack.
+    In the development of the Internet, the TCP protocol that has a very special place.
+    Initially described by the ``fathers of the Internet'' Vincent Cerf and Bob Kahn, the TCP protocol was standardized in 1981 by RFC 793~\cite{rfc793}.
+    Since this date, TCP has been widely used on the Internet as the underlying protocol for application layers such as HTTP,
+    FTP or TLS. TCP is based on the underlying IP protocol which is itself
+    based on a link layer such as Ethernet. This stacking of protocols is referred to as the TCP/IP stack.
 
-    The last decade has seen a growing interest for security and privacy in all the fields where computers are
-    involved. The development of safe applications thanks to static analyzers and formal methods, that was first
-    restricted to critical embedded software with the Astr\'ee analyzer~\cite{cousot2005astree} or the B method, are
-    now adopted on a larger scale. New tools have been developed to promote the use of formal methods. Among them we
-    can mention the SMT-solvers Z3~\cite{de2008z3} and Alt-Ergo~\cite{conchon:hal-01960203},
-    the higher level language F*~\cite{mumon} that can prove assertions
-    using these SMT-solvers or the more automatic static analyzer by abstract interpretation Frama-C~\cite{cuoq2012frama}
-    that can help to identify undefined behaviors in C programs. Theorem provers, such as Coq or Isabelle, are another class
-    of tools, where proofs are done manually.
+    Given the importance of the TCP protocol in network communications, some researchers have attempted to formalize the protocol to either prove key properties~\cite{smith1996formal}
+    or even to fully prove the correctness of an implementation against a complete specification~\cite{ridge2008rigorous}.
+    This latter implementation was performed in the HOL proof assistant, and while the authors note that it would be possible in theory to extract an implementation in Haskell from
+    this work, this has not been adopted in practice.
+    The formalization of TCP specification and verification of a TCP implementation face two difficulties:
+    RFC 793 is written in English, leaving many parts underspecified; the TCP protocol is inherently concurrent which makes it complex to specify and verify.
+    As a result, there is currently no formally verified implementation for TCP that is usable in industry.
 
-    In the domain of networking and cryptography, these tools have been used to make secure implementations.
-    Recently, Reiher and al have proposed a domain-specific language, RecordFlux~\cite{Reiher2019RecordFluxFM},
-    to specify the message formats for real world protocols, that is written in SPARK.
-    This is also the goal followed by the Everest project~\cite{bhargavan2017everest}.
-    The authors implemented cryptographic functions such as AES or
-    SHA2 as well as TLS, the cryptographic protocol that provides security
-    to HTTPS and is its heart. The implementation is written
-    in F* and is a combination of several sub-project such as miTLS~\cite{bhargavan2013implementing}
-    or HACL*~\cite{zinzindohoue2017hacl}. The weakness of the Everest project lies in its need to use the untrusted
-    transport layer TCP. If a bug exists in the transport layer, it can also jeopardize the higher level layers.
+    Other communication protocols have been formalized in recent years, leading to a formally verified implementation.
+    The most prominent project in this area is the Everest project~\cite{bhargavan2017everest}, which led to a formally verified implementation in F* of TLS called miTLS~\cite{bhargavan2013implementing}.
+    Note that miTLS still relies on an untrusted implementation of TCP, and thus could benefit from the results of our work.
+    The approach taken in project Everest was to reimplement the formalized protocols completely in F*, a research programming language targeted at formal program verification.
+    This approach is often difficult to adopt in industry, as codebases are evolved incrementally rather than replaced, both for economical and practical reasons.
 
-    The importance of the TCP protocol in the network communications has driven some people to work
-    on it and try to ensure the safety of the protocol. Among these works, in~\cite{smith1996formal}
-    the author has tried to formally verify that a TCP implementation verify some specifications.
-    Another work comes from Tom Ridge and al~\cite{ridge2008rigorous} who have implemented the TCP
-    protocol in the HOL proof assistant.
-    But only few works has been done. Doing formal verification on a TCP implementation face up to two
-    problems. The first one is the complexity of the TCP protocol that is written to be concurrent.
-    The second problem comes from the norm itself that is written in a natural, informal language: English.
-    Then a lot of part are under specified and extracted strong specifications can be hard.
-    As a result, it does not exists a safe implementation for TCP that can be used in the industry.
-
-    Our approach has been to consider an existing implementation of a TCP/IP stack and replace and integrate
-    some bricks by Ada/SPARK code.
-    We have chosen the Oryx Embedded's library, CycloneTCP\footnote{\url{https://www.oryx-embedded.com/products/CycloneTCP.html}}.
-    The library is written in C, which is suitable to interface with Ada/SPARK, and is developed to run
-    on embedded processors. We focused our work on rewriting in SPARK the TCP protocol because of the importance
-    it has in the network communication as well as the socket API to provide a safe interface to the programmers
-    to use our safe TCP implementation.
-
-    In the first section we explore more in details the TCP protocol and the state machine that governs the behavior of the connection.
-    Then in the next section we explain how we have formalized and implemented in SPARK the TCP user functions and the socket API.
-    Finally in the last section we present some results of our work, like the bugs that have been found and eliminated thanks to our
-    new implementation.
+    Our approach in this work has been to incrementally evolve an existing TCP/IP stack in C by replacing parts of the code with formally verified code in SPARK.
+    The stack chosen, CycloneTCP\footnote{\url{https://www.oryx-embedded.com/products/CycloneTCP.html}}, is targeted at embedded processors with no Operating System.
+    In Section~\ref{sec:TCP} we describe the TCP protocol, in particular the state machine that governs the behavior of a connection.
+    In Section~\ref{sec:spec} we present specificities of the specification technique that we used, and in Section~\ref{sec:verif} we explain how we formalized and implemented in SPARK the TCP user functions and the underlying socket API.
+    In Section~\ref{sec:results} we present the results of our work, in particular bugs that have been found and eliminated thanks to our approach, before we conclude.
 
@@
 \section{Specification Techniques Used}
+\label{sec:spec}
 
@@
 \section{Verification of the TCP protocol and the Socket API}
+\label{sec:verif}
 
@@
 \section{Results and bugs found}
+\label{sec:results}
 
